---
title: "Empirical use-resistance relationships over geographic scales"
author: "Scott Olesen"
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(
  fig.width = 7, fig.height = 4, fig.path = 'fig/',
  dev = c('png', 'pdf'),
  echo = FALSE, warning = FALSE, message = FALSE,
  cache = TRUE, autodep = TRUE)

pdf.options(useDingbats = FALSE, useKerning = FALSE)

library(MASS)
map = purrr::map
select = dplyr::select
```

# Methods

## Use and resistance data

### MarketScan

```{r maup_db}
state_data = read_tsv('../db/state_data.tsv') %>%
  rename(us_region = region, us_division = division) %>%
  mutate(density = population / area)
```

```{r marketscan_data}
marketscan_res = read_tsv('~/grad/proj/medicare/analysis/ms2/data/abg_state.tsv', col_types = cols('n_isolates' = 'n')) %>%
  # specify n_isolates as "n" to parse things like "5e3"
  mutate_at('n_isolates', as.integer) %>%
  rename(drug = drug_group) %>%
  mutate(bugdrug = case_when(
    .$bug == 'E. coli' & .$drug == 'quinolone' ~ 'Ec/q',
    .$bug == 'S. pneumoniae' & .$drug == 'beta_lactam' ~ 'Sp/bl',
    .$bug == 'S. pneumoniae' & .$drug == 'macrolide' ~ 'Sp/m'
  )) %>%
  filter(!is.na(bugdrug)) %>%
  mutate(n_resistant = as.integer(round(f_ns * n_isolates))) %>%
  select(bugdrug, drug, state, n_resistant, n_isolates)

marketscan_use = read_tsv('~/grad/proj/medicare/analysis/ms2/ineq_marketscan.tsv') %>%
  rename(drug = drug_group, use = total_use) %>%
  filter(drug %in% c('quinolone', 'beta_lactam', 'macrolide'))

marketscan = marketscan_use %>%
  inner_join(marketscan_res, by = c('drug', 'state')) %>%
  rename(unit = state) %>%
  left_join(state_data, by = 'unit') %>%
  select(
    unit, bugdrug, use, n_resistant, n_isolates,
    population, density, income, temperature,
    us_region, us_division
  )
```

### NHSN/IMS

```{r nhsn_ims}
nhsn = read_tsv('../data/nhsn-ims/data.tsv') %>%
  rename(state_abbreviation = state) %>%
  filter(state_abbreviation %in% state.abb) %>%
  mutate(
    unit_id = match(state_abbreviation, state.abb),
    bugdrug = 'Ec/q',
    use = rx_1k_year / 1e3
  ) %>%
  left_join(state_data, by = 'unit_id') %>%
  select(
    unit, bugdrug, use, n_resistant, n_isolates,
    population, density, income, temperature,
    us_region, us_division
  )
```

### ECDC

For the European data, we convert to "treatments" per person per year, assuming a certain conversion from DDD to treatments. This is just an *ad hoc* adjustment to get the European data to the same scales as used in the simulations and the US data.

```{r ecdc_convert}
did_cpy_map = data_frame(
  drug = c('beta_lactam', 'quinolone', 'macrolide'),
  ddd_per_tx = c(10, 10, 7),
  cpy_per_did = 365 / (1e3 * ddd_per_tx)
)

did_cpy_map %>%
  kable(caption = 'DDD to treatment conversion')
```

```{r ecdc}
europe_units = read_tsv('../db/europe_units.tsv') %>%
  mutate(density = population / area) %>%
  select(unit_id, unit, population, eu_region = region, density, income, temperature)

europe = read_tsv('../data/ecdc/data.tsv') %>%
  left_join(did_cpy_map, by = 'drug') %>%
  mutate(use = cpy_per_did * did) %>%
  mutate(bugdrug = case_when(
    .$bug == 'Escherichia coli' & .$drug == 'quinolone' ~ 'Ec/q',
    .$bug == 'Streptococcus pneumoniae' & .$drug == 'beta_lactam' ~ 'Sp/bl',
    .$bug == 'Streptococcus pneumoniae' & .$drug == 'macrolide' ~ 'Sp/m'
  )) %>%
  rename(unit = country) %>%
  left_join(europe_units, by = 'unit') %>%
  select(
    unit, bugdrug, use, n_resistant = n_ns, n_isolates,
    population, density, temperature, income,
    eu_region
  )

europe_names = unique(europe$unit)
```

```{r combine_data}
named_bind_rows = function(..., label = 'dataset') {
  names = map_chr(quos(...), quo_name)
  map2_dfr(list(...), names, ~ mutate(.x, !!as.symbol(label) := .y))
}

unit_data = named_bind_rows(marketscan, nhsn, europe) %>%
  mutate(
    n_susceptible = n_isolates - n_resistant,
    f_resistant = n_resistant / n_isolates
  )

# look for dataset/region combinations to do
region_combinations = unit_data %>%
  group_by(dataset, bugdrug) %>%
  summarize_at(c('unit', 'us_region', 'us_division', 'eu_region'), ~ not(any(is.na(.)))) %>%
  ungroup() %>%
  gather('group_type', 'is_present', -dataset, -bugdrug) %>%
  filter(is_present) %>% select(-is_present) %>%
  arrange(dataset, bugdrug, group_type)
```

## Adjacency

```{r adjacency_db}
us_adjacency = read_tsv('../db/state_adjacency.tsv')
eu_adjacency = read_tsv('../db/europe_adjacency.tsv')
```

# Results

## Regional analysis

Aggregate the data into regions: Census regions and divisions for US data, and UN Geoscheme regions for European data. Aggregate use data by population-weighted means. Aggregate resistance data by adding up the numbers of susceptible and resistant isolates.

```{r region_data}
aggregate_at = function(df, group_chr) {
  stopifnot(class(group_chr) == 'character')
  group = as.symbol(group_chr)
  
  group_by_at(df, group_chr, .add = TRUE) %>%
    summarize(
      use = weighted.mean(use, w = population),
      n_resistant = sum(n_resistant),
      n_isolates = sum(n_isolates)
    ) %>%
    ungroup() %>%
    rename(group := !!group) %>%
    mutate(
      group_type = group_chr,
      n_susceptible = n_isolates - n_resistant
    )
}

region_data = region_combinations %>%
  mutate(
    unit_data = map2(
      dataset, bugdrug,
      ~ filter(unit_data, dataset == .x, bugdrug == .y)
    ),
    region_data = map2(unit_data, group_type, aggregate_at)
  )
```

Run logistic regressions predicting resistance from use rates for each dataset, bug/drug, and group type. Group types include baseline "units" (US states or European countries), US Census regions, etc. Get a $p$-value on whether the slopes in the aggregates are different than the ones in the baseline units by shuffling the region labels.

```{r regional_analysis}
model_f = function(df) glm(
  cbind(n_resistant, n_susceptible) ~ use,
  family = 'binomial', data = df
)

# compute indices, models, and predictions
region_boot_f = function(df, group_chr, n_trials = 1e3, alpha = 0.05, pred_n = 1e4) {
  pred_use = seq(min(df$use), max(df$use), length.out = pred_n)
  newdata = data_frame(use = pred_use)
  
  # get predictions from base model
  model = df %>%
    aggregate_at(group_chr) %>%
    model_f

  model_pred = data_frame(
    use = pred_use,
    res = predict(model, newdata = newdata, type = 'response')
  )
  
  results = data_frame(i = 1:n_trials) %>%
    mutate(
      # get bootstrapped data sets
      unit_data = map(i, ~ sample_n(df, size = nrow(df), replace = TRUE)),
      data = map(unit_data, ~ aggregate_at(., group_chr)),
      # run the model and predict the outcomes
      model = map(data, model_f),
      slope = map_dbl(model, ~ coef(.)['use']),
      pred_res = map(model, ~ predict(., newdata = newdata, type = 'response')),
      pred_df = map(pred_res, ~ data_frame(use = pred_use, res = .))
    )
  
  # at each point pred_use, get the CIs for the predictions
  pred_data = results %>%
    select(i, pred_df) %>% unnest() %>%
    group_by(use) %>%
    summarize(
      ymin = quantile(res, alpha / 2),
      ymax = quantile(res, 1 - alpha / 2)
    ) %>%
    left_join(model_pred, by = 'use')
  
  list(
    slope_cil = quantile(results$slope, alpha / 2),
    slope_ciu = quantile(results$slope, 1 - alpha / 2),
    pred_data = pred_data
  )
}

region_models = region_data %>%
  mutate(
    # fit models to each dataset, bugdrug, & aggregate type
    model = map(region_data, model_f),
    # extract "slope" from base model
    slope = map_dbl(model, ~ coef(.)['use']),
    # bootstrapping
    boot_results = map2(unit_data, group_type, region_boot_f),
    slope_cil = map_dbl(boot_results, ~ .$slope_cil),
    slope_ciu = map_dbl(boot_results, ~ .$slope_ciu),
    pred_data = map(boot_results, ~ .$pred_data)
  ) %>%
  # the "base" slopes are the cross-unit measurements
  group_by(dataset, bugdrug) %>%
  mutate(base_slope = slope[group_type == 'unit']) %>%
  ungroup() %>%
  arrange(dataset, bugdrug, group_type)
```

### Plots

```{r regional_results_plot}
my_palette = c('black', '#1b9e77', '#d95f02')

plotify_region_data = function(df) {
  df %>%
    mutate(
      dataset = recode(dataset, 'europe' = 'ECDC', 'marketscan' = 'MarketScan/RO', 'nhsn' = 'IMS/NHSN'),
      long_bugdrug = recode(bugdrug, 'Ec/q' = 'E. coli & quinolones', 'Sp/m' = 'S. pneumonia & macrolides', 'Sp/bl' = 'S. pneumoniae & beta-lactams'),
      label = str_c(dataset, ' ', bugdrug),
      group_fct = factor(case_when(
        group_type == 'unit' ~ 'unit',
        group_type == 'us_division' ~ 'division',
        group_type %in% c('us_region', 'eu_region') ~ 'region',
        TRUE ~ NA_character_
      ), levels = c('unit', 'division', 'region'))
    )
}

# pull out the data and the predictions separately
region_plot_data = region_models %>%
  select(dataset, bugdrug, group_type, region_data) %>%
  unnest() %>%
  mutate(res = n_resistant / n_isolates) %>%
  plotify_region_data()

region_pred_data = region_models %>%
  select(dataset, bugdrug, group_type, pred_data) %>%
  unnest() %>%
  plotify_region_data()

# and just the marketscan data, for the main text figure
marketscan_region_plot_data = region_plot_data %>%
  filter(dataset == 'MarketScan/RO')

marketscan_region_pred_data = region_pred_data %>%
  filter(dataset == 'MarketScan/RO')

marketscan_regional_plot = ggplot(data = NULL, aes(use, res, color = group_fct, fill = group_fct)) +
  facet_wrap(~ long_bugdrug, scales = 'free') +
  geom_ribbon(
    data = marketscan_region_pred_data,
    aes(ymin = ymin, ymax = ymax, color = NULL),
    alpha = 0.25
  ) +
  geom_line(data = marketscan_region_pred_data, size = 1) +
  geom_point(data = marketscan_region_plot_data) +
  scale_fill_manual(values = my_palette) +
  scale_color_manual(values = my_palette) +
  xlab('antibiotic use (annual treatments per capita)') +
  scale_y_continuous(
    'antibiotic resistance (% of isolates)',
    labels = scales::percent_format(accuracy = 1, suffix = '')
  ) +
  theme_classic() +
  theme(
    legend.position = 'none',
    strip.text = element_text(size = 8),
    strip.background = element_blank(),
    axis.text = element_text(size = 8, color = 'black'),
    axis.title = element_text(size = 8)
  )

show(marketscan_regional_plot)

ggsave(
  'fig/marketscan_regional.pdf',
  plot = marketscan_regional_plot,
  width = 180, height = 80, units = 'mm'
)
```

```{r other_regional_results_plot}
# and the other stuff
other_region_plot_data = region_plot_data %>%
  filter(dataset != 'MarketScan/RO')

other_region_pred_data = region_pred_data %>%
  filter(dataset != 'MarketScan/RO')

other_regional_plot = ggplot(data = NULL, aes(use, res, color = group_fct, fill = group_fct)) +
  facet_wrap(~ label, scales = 'free') +
  geom_ribbon(
    data = other_region_pred_data,
    aes(ymin = ymin, ymax = ymax, color = NULL),
    alpha = 0.25
  ) +
  scale_fill_manual(values = my_palette) +
  scale_color_manual(values = my_palette) +
  geom_point(data = other_region_plot_data) +
  geom_line(data = other_region_pred_data) +
  xlab('antibiotic use (annual treatments per capita)') +
  scale_y_continuous(
    'antibiotic resistance (% of isolates)',
    labels = scales::percent_format(accuracy = 1, suffix = '')
  ) +
  theme_classic() +
  theme(
    legend.position = 'none',
    strip.text = element_text(size = 8),
    strip.background = element_blank(),
    axis.text = element_text(size = 8, color = 'black'),
    axis.title = element_text(size = 8)
  )

show(other_regional_plot)

ggsave(
  'fig/other_regional.pdf',
  plot = other_regional_plot,
  width = 180, height = 120, units = 'mm'
)
```

```{r regional_slopes_plot}
region_slopes_plot = region_models %>%
  mutate(
    dataset = recode(dataset, 'europe' = 'ECDC', 'marketscan' = 'MarketScan/RO', 'nhsn' = 'IMS/NHSN'),
    group_type = factor(
      group_type,
      levels = c('unit', 'us_division', 'us_region', 'eu_region'),
      labels = c('Unit', 'Division', 'Region', 'Region')
    ),
    long_bugdrug = recode(bugdrug, 'Ec/q' = 'E. coli & quinolones', 'Sp/m' = 'S. pneumonia & macrolides', 'Sp/bl' = 'S. pneumoniae & beta-lactams'),
    label = str_c(dataset, ' ', bugdrug)
  ) %>%
  ggplot(aes(group_type, slope, color = group_type, ymin = slope_cil, ymax = slope_ciu)) +
  facet_wrap(~ label, scales = 'free') +
  geom_point() +
  geom_errorbar() +
  # make US and European regions the same color
  scale_color_manual(values = my_palette[c(1, 2, 3, 3)]) +
  xlab('') +
  ylab('Regression coefficient') +
  theme_classic() +
  theme(
    legend.position = 'none',
    strip.text = element_text(size = 8),
    strip.background = element_blank(),
    axis.text = element_text(size = 8, color = 'black'),
    axis.title = element_text(size = 8)
  )

show(region_slopes_plot)

ggsave(
  'fig/region_slopes.pdf',
  plot = region_slopes_plot,
  width = 180, height = 120, units = 'mm'
)
```

## Distance analysis

If local transmission "washes out" the use/resistance relationship, then we expect that adjacent geographic units will have a *smaller* difference in resistance for the same difference in use. (Consider the extreme: two perfectly-interacting populations should have almost no difference in resistance, regardless of their difference in use.) That is, in a regression
$$
\Delta \rho \sim \Delta \tau + \Delta \tau \times a + \Delta \mathrm{income} + \ldots
$$
where $a$ is a flag for adjacency, we expect the interaction term $\beta_{\tau a}$ will be negative.

```{r distance_analysis}
# because no US states and European countries share names, it's OK to combine this list
adjacency_db = bind_rows(us_adjacency, eu_adjacency) %>%
  mutate(adjacent = 1)

odds = function(p) p / (1 - p)
log_odds_ratio = function(p, q) log(odds(p)) - log(odds(q))

cross_units = function(df) {
  df %$%
    crossing(unit1 = unit, unit2 = unit) %>%
    left_join(rename_all(df, ~ str_c(., '1')), by = 'unit1') %>%
    left_join(rename_all(df, ~ str_c(., '2')), by = 'unit2') %>%
    filter(use1 > use2) %>%
    mutate(
      d_resistant = f_resistant1 - f_resistant2,
      d_use = log_odds_ratio(use1, use2),
      d_income = income1 - income2,
      d_density = density1 - density2,
      d_temperature = temperature1 - temperature2
    ) %>%
    select(unit1, unit2, starts_with('d_')) %>%
    left_join(adjacency_db, by = c('unit1', 'unit2')) %>%
    # pairs of units not in the adjacency list are not adjacent
    replace_na(list(adjacent = 0))
}

bisquare = partial(MASS::rlm, psi = MASS::psi.bisquare)
pair_model_f = function(df) bisquare(d_resistant ~ d_use + d_use:adjacent + d_income + d_density + d_temperature, data = df)

# compute indices, models, and predictions
distance_boot_f = function(df, n_trials = 1e3, alpha = 0.05, pred_n = 1e4) {
  base_x = df %>% cross_units %$% d_use
  pred_d_use = seq(min(base_x), max(base_x), length.out = pred_n)
  newdata = crossing(
    d_use = pred_d_use,
    adjacent = c(0, 1),
    d_income = 0,
    d_density = 0,
    d_temperature = 0
  )
  
  results = data_frame(i = 1:n_trials) %>%
    mutate(
      # get bootstrapped data sets
      unit_data = map(i, ~ sample_n(df, size = nrow(df), replace = TRUE)),
      cross_data = map(unit_data, cross_units),
      # run the model and predict the outcomes
      model = map(cross_data, pair_model_f),
      inter = map_dbl(model, ~ coef(.)['d_use:adjacent']),
      pred_y = map(model, ~ predict(., newdata = newdata, type = 'response')),
      pred_df = map(pred_y, ~ mutate(newdata, d_res = .))
    )
  
  # at each point pred_use, get the CIs for the predictions
  pred_data = results %>%
    select(i, pred_df) %>% unnest() %>%
    group_by(d_use, adjacent) %>%
    summarize(
      ymin = quantile(d_res, alpha / 2),
      ymax = quantile(d_res, 1 - alpha / 2)
    ) %>%
    ungroup()
  
  list(
    inter_cil = quantile(results$inter, alpha / 2),
    inter_ciu = quantile(results$inter, 1 - alpha / 2),
    pred_data = pred_data
  )
}

distance_results = unit_data %>%
  nest(-dataset, -bugdrug) %>%
  mutate(
    pair_data = map(data, cross_units),
    model = map(pair_data, pair_model_f),
    inter = map_dbl(model, ~ coef(.)['d_use:adjacent']),
    # bootstrapping
    boot_results = map(data, distance_boot_f),
    inter_cil = map_dbl(boot_results, ~ .$inter_cil),
    inter_ciu = map_dbl(boot_results, ~ .$inter_ciu),
    pred_data = map(boot_results, ~ .$pred_data)
  )
```

```{r distance_results_table}
distance_results %>%
  mutate(
    dataset = recode(dataset, 'europe' = 'ECDC', 'marketscan' = 'MarketScan/RO', 'nhsn' = 'IMS/NHSN'),
    interaction = sprintf('%.2f (%.2f to %.2f)', inter, inter_cil, inter_ciu)
  ) %>%
  select(dataset, bugdrug, interaction) %>%
  arrange(dataset, bugdrug) %>%
  kable()
```

### Plots

First plot shows MS/RO and Sp/m.

```{r distance_plots}
main_distance_plot_data = distance_results %>%
  filter(dataset == 'marketscan', bugdrug == 'Sp/m') %$%
  pair_data[[1]] %>%
  mutate(adjacent = adjacent == 1)

main_distance_pred_data = distance_results %>%
  filter(dataset == 'marketscan', bugdrug == 'Sp/m') %$%
  pred_data[[1]] %>%
  mutate(adjacent = adjacent == 1)

main_distance_plot = ggplot(data = NULL, aes(d_use, color = adjacent, fill = adjacent)) +
  geom_point(data = main_distance_plot_data, aes(y = d_resistant), shape = 1) +
  stat_smooth(data = main_distance_plot_data, aes(y = d_resistant), method = 'rlm', se = FALSE) +
  geom_ribbon(
    data = main_distance_pred_data,
    aes(ymin = ymin, ymax = ymax, color = NULL),
    alpha = 0.25
  ) + 
  scale_color_manual(values = c('black', 'red')) +
  scale_fill_manual(values = c('black', 'red')) +
  xlab('Difference in antibiotic use (annual treatments per capita)') +
  ylab('Difference in resistance (log odds ratio)') +
  theme_classic() +
  theme(
    legend.position = 'none',
    axis.text = element_text(size = 8, color = 'black'),
    axis.title = element_text(size = 8)
  )

show(main_distance_plot)

other_distance_plot_data = distance_results %>%
  filter(!(dataset == 'marketscan' & bugdrug == 'Sp/m')) %>%
  select(dataset, bugdrug, pair_data) %>%
  unnest() %>%
  mutate(
    adjacent = adjacent == 1,
    dataset = recode(dataset, 'europe' = 'ECDC', 'marketscan' = 'MarketScan/RO', 'nhsn' = 'IMS/NHSN'),
    label = str_c(dataset, ' ', bugdrug)
  )

other_distance_pred_data = distance_results %>%
  filter(!(dataset == 'marketscan' & bugdrug == 'Sp/m')) %>%
  select(dataset, bugdrug, pred_data) %>%
  unnest() %>%
  mutate(
    adjacent = adjacent == 1,
    dataset = recode(dataset, 'europe' = 'ECDC', 'marketscan' = 'MarketScan/RO', 'nhsn' = 'IMS/NHSN'),
    label = str_c(dataset, ' ', bugdrug)
  )

other_distance_plot = ggplot(
  data = NULL, aes(d_use, color = adjacent, fill = adjacent)
) +
  facet_wrap(~ label, scales = 'free') +
  geom_point(
    data = other_distance_plot_data, aes(y = d_resistant),
    shape = 1, size = 0.5
  ) +
  stat_smooth(
    data = other_distance_plot_data,
    aes(y = d_resistant), method = 'rlm', se = FALSE
  ) +
  geom_ribbon(
    data = other_distance_pred_data,
    aes(ymin = ymin, ymax = ymax, color = NULL), alpha = 0.25
  ) + 
  scale_color_manual(values = c('black', 'red')) +
  scale_fill_manual(values = c('black', 'red')) +
  xlab('Difference in antibiotic use (annual treatments per capita)') +
  ylab('Difference in resistance (log odds ratio)') +
  theme_classic() +
  theme(
    legend.position = 'none',
    strip.text = element_text(size = 8),
    strip.background = element_blank(),
    axis.text = element_text(size = 8, color = 'black'),
    axis.title = element_text(size = 8)
  )

show(other_distance_plot)

ggsave(
  'fig/main_distance_plot.pdf',
  plot = main_distance_plot,
  width = 88, height = 80, units = 'mm'
)

ggsave(
  'fig/other_distance_plot.pdf',
  plot = other_distance_plot,
  width = 180, height = 80, units = 'mm'
)
```
