---
title: "Geography and associations between antibiotic use and resistance: observational data"
author: "Scott Olesen"
---

```{r global_options, echo = FALSE, message = FALSE}
library(knitr)
library(cowplot)
library(tidyverse)

knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE,
  cache = TRUE, autodep = TRUE
)

pdf.options(useDingbats = FALSE, useKerning = FALSE)
```

# Load data

```{r state_data}
us_temp <- read_tsv('../db/us/temperature.tsv')
us_income <- read_tsv('../db/us/income.tsv')
us_density <- read_tsv('../db/us/density.tsv')

us_state_data <- us_temp %>%
  left_join(us_income, by = 'state') %>%
  left_join(us_density, by = 'state')
```

```{r marketscan_data}
marketscan_res <- read_tsv('../data/ms-medicare-ro/abg_state.tsv') %>%
  rename(drug = drug_group) %>%
  mutate(bugdrug = case_when(
    # don't do Sp/bl, as they have negative correlation
    .$bug == 'E. coli' & .$drug == 'quinolone' ~ 'Ec/q',
    .$bug == 'S. pneumoniae' & .$drug == 'macrolide' ~ 'Sp/m'
  )) %>%
  filter(!is.na(bugdrug)) %>%
  select(bugdrug, drug, state, f_resistant = f_ns)

marketscan_use <- read_tsv('../data/ms-medicare-ro/ineq_marketscan.tsv') %>%
  rename(drug = drug_group, use = total_use) %>%
  filter(drug %in% c('quinolone', 'beta_lactam', 'macrolide'))

marketscan <- marketscan_use %>%
  inner_join(marketscan_res, by = c('drug', 'state')) %>%
  left_join(us_state_data, by = 'state') %>%
  select(
    unit = state, bugdrug, use, f_resistant,
    density, income, temperature
  )
```

```{r nhsn_ims}
nhsn <- read_tsv('../data/nhsn-ims/data.tsv') %>%
  rename(state_abbreviation = state) %>%
  filter(state_abbreviation %in% state.abb) %>%
  mutate(
    state = state.name[match(state_abbreviation, state.abb)],
    bugdrug = 'Ec/q',
    use = rx_1k_year / 1e3,
    f_resistant = n_resistant / n_isolates
  ) %>%
  left_join(us_state_data, by = 'state') %>%
  select(
    unit = state, bugdrug, use, f_resistant,
    density, income, temperature
  )
```

For the European data, we convert to "treatments" per person per year, assuming a certain conversion from DDD to treatments. This is just an *ad hoc* adjustment to get the European data to the same scales as used in the simulations and the US data.

```{r ecdc_convert}
did_cpy_map <- tibble(
  drug = c('beta_lactam', 'quinolone', 'macrolide'),
  ddd_per_tx = c(10, 10, 7),
  cpy_per_did = 365 / (1e3 * ddd_per_tx)
)

did_cpy_map %>%
  kable(caption = 'DDD to treatment conversion')
```

```{r ecdc}
eu_temp <- read_tsv('../db/europe/temperature.tsv')
eu_income <- read_tsv('../db/europe/income.tsv')
eu_density <- read_tsv('../db/europe/density.tsv')

europe <- read_tsv('../data/ecdc/data.tsv') %>%
  left_join(did_cpy_map, by = 'drug') %>%
  mutate(use = cpy_per_did * did) %>%
  mutate(
    bugdrug = case_when(
      .$bug == 'Escherichia coli' & .$drug == 'quinolone' ~ 'Ec/q',
      .$bug == 'Streptococcus pneumoniae' & .$drug == 'beta_lactam' ~ 'Sp/bl',
      .$bug == 'Streptococcus pneumoniae' & .$drug == 'macrolide' ~ 'Sp/m'
    ),
    f_resistant = n_ns / n_isolates
  ) %>%
  left_join(eu_temp, by = 'country') %>%
  left_join(eu_income, by = 'country') %>%
  left_join(eu_density, by = 'country') %>%
  rename(unit = country) %>%
  select(
    unit, bugdrug, use, f_resistant,
    density, temperature, income
  )

eu_units <- unique(europe$unit)
```

```{r combine_data}
unit_data <- bind_rows(
  'MarketScan/RO' = marketscan,
  'Xponent/NHSN' = nhsn,
  'ECDC' = europe,
  .id = 'data_source'
) %>%
  mutate(dataset = str_c(data_source, ' ', bugdrug)) %>%
  select(dataset, unit, f_resistant, use, density, temperature, income) %>%
  nest(-dataset) %>%
  mutate_at('dataset', fct_inorder)
```

```{r adjacency_data}
us_adjacency <- read_tsv('../db/us/adjacency.tsv') %>%
  mutate(adjacent = TRUE) %>%
  complete(state1, state2) %>%
  replace_na(list(adjacent = FALSE)) %>%
  rename(unit1 = state1, unit2 = state2)

eu_adjacency <- read_tsv('../db/europe/adjacency.tsv') %>%
  mutate(adjacent = TRUE) %>%
  right_join(crossing(country1 = eu_units, country2 = eu_units)) %>%
  replace_na(list(adjacent = FALSE)) %>%
  rename(unit1 = country1, unit2 = country2)

adjacency_db <- bind_rows(us_adjacency, eu_adjacency)
```

```{r commuting_data}
us_commuting <- read_tsv('../db/us/commuting.tsv') %>%
  rename_all(~ str_replace(., '^state', 'unit'))

eu_commuting <- read_tsv('../db/europe/commuting.tsv') %>%
  rename_all(~ str_replace(., '^country', 'unit')) %>%
  filter(unit1 %in% eu_units, unit2 %in% eu_units)

commuting_db <- bind_rows(
  'US' = us_commuting,
  'Europe' = eu_commuting,
  .id = 'dataset'
) %>%
  mutate_at('dataset', fct_inorder)
```

# Use-resistance in different datasets

```{r use_resistance_plots}
round_up <- function(x, digits) ceiling(x * 10 ** digits) / 10 ** digits
round_down <- function(x, digits) floor(x * 10 ** digits) / 10 ** digits

breaker <- function(digits) {
  function(x) {
    lower <- round_up(x[1], digits)
    upper <- round_down(x[2], digits)
    round(seq(lower, upper, length.out = 3), digits)
  }
}

obs_plot <- unit_data %>%
  unnest() %>%
  ggplot(aes(use, f_resistant * 100)) +
  facet_wrap(~ dataset, scales = 'free') +
  geom_smooth(method = 'lm', color = 'gray50') +
  geom_point() +
  scale_x_continuous(
    'Treatments per person per year',
    breaks = breaker(2)
  ) +
  scale_y_continuous(
    name = 'Resistance (% isolates nonsusceptible)'
  ) +
  theme_cowplot() +
  theme(
    strip.background = element_blank(),
    plot.margin = margin(1, 5, 1, 1, 'mm')
  )

show(obs_plot)
ggsave('fig/obs_plot.pdf', plot = obs_plot)
ggsave('fig/obs_plot.png', plot = obs_plot)
```

# Pairs data

```{r cross_data}
odds <- function(p) p / (1 - p)
log_odds_ratio <- function(p, q) log(odds(p)) - log(odds(q))

cross_units <- function(df) {
  df %>%
    with(crossing(unit1 = unit, unit2 = unit)) %>%
    filter(unit1 < unit2) %>%
    left_join(rename_all(df, ~ str_c(., '1')), by = 'unit1') %>%
    left_join(rename_all(df, ~ str_c(., '2')), by = 'unit2') %>%
    mutate(
      d_resistant = f_resistant1 - f_resistant2,
      d_use = use1 - use2,
      d_income = income1 - income2,
      d_density = density1 - density2,
      d_temperature = temperature1 - temperature2,
      dr_du = d_resistant / d_use,
      lor_resistant = log_odds_ratio(f_resistant1, f_resistant2),
      lorr_du = lor_resistant / d_use
    ) %>%
    select(unit1, unit2, starts_with('d_'), dr_du, starts_with('lor')) %>%
    left_join(adjacency_db, by = c('unit1', 'unit2')) %>%
    left_join(commuting_db, by = c('unit1', 'unit2'))
}

leave_one_out_from_cross <- function(df) {
  units <- unique(c(df$unit1, df$unit2))
  map(units, ~ filter(df, unit1 != ., unit2 != .))
}

cross_data <- unit_data %>%
  mutate(
    cross_data = map(data, cross_units),
    l1o_cross_data = map(cross_data, leave_one_out_from_cross)
  )
```

## Commuting histogram

```{r commuting_histogram}
commuting_histogram <- cross_data %>%
  select(cross_data) %>%
  unnest() %>%
  select(dataset, unit1, unit2, adjacent, f_commuting) %>%
  filter(
    unit1 < unit2,
    !is.na(f_commuting)
  ) %>%
  ggplot(aes(log10(f_commuting), fill = adjacent)) +
  facet_wrap(~ dataset, scales = 'free') +
  geom_histogram(color = 'black') +
  scale_fill_manual(
    limits = c('TRUE', 'FALSE'),
    labels = c('adjacent', 'not adj.'),
    values = c('black', 'white')
  ) +
  xlab('Commuting fraction (log 10)') +
  ylab('Number of states or countries') +
  theme_cowplot() +
  theme(
    strip.background = element_blank(),
    legend.position = c(0.85, 0.75),
    legend.title = element_blank()
  )

show(commuting_histogram)
ggsave('fig/commuting_histogram.pdf', plot = commuting_histogram)
ggsave('fig/commuting_histogram.png', plot = commuting_histogram)
```

Some of the pairs had zero interactions and don't appear in this figure:

```{r}
commuting_db %>%
  filter(unit1 < unit2, !is.na(f_commuting)) %>%
  count(dataset, f_commuting == 0) %>%
  kable(caption = 'Number of pairs with 0 commuting fraction')
```

```{r analysis_helpers}
sigfig <- function(x, n = 2) {
  formatC(signif(x, digits = n), digits = n, format = "fg", flag = "#")
}

jackknife.sd <- function(x) {
  n <- length(x)
  sqrt((n - 1) / n * sum((x - mean(x)) ** 2))
}

analysis_f <- function(model_f, coef_f, ratio_f) {
  base_results <- cross_data %>%
    mutate(
      model = map(cross_data, model_f),
      coef = map_dbl(model, coef_f),
      ratio = map_dbl(model, ratio_f)
    ) %>%
    select(dataset, coef, ratio)
  
  l1o_results <- cross_data %>%
    select(dataset, l1o_cross_data) %>%
    unnest() %>%
    mutate(
      model = map(l1o_cross_data, model_f),
      coef = map_dbl(model, coef_f),
      ratio = map_dbl(model, ratio_f)
    ) %>%
    group_by(dataset) %>%
    summarize(
      coef_se = jackknife.sd(coef),
      ratio_se = jackknife.sd(ratio)
    )
  
  base_results %>%
    left_join(l1o_results, by = 'dataset')
}

rlm <- MASS::rlm
```

```{r adjacency_analysis}
adjacency_results <- analysis_f(
  function(df) with(df, {
    list(
      adj_med = median(df$dr_du[df$adjacent]),
      nonadj_med = median(df$dr_du[!df$adjacent])
    )
  }),
  function(model) with(model, { adj_med - nonadj_med }),
  function(model) with(model, { (adj_med - nonadj_med) / nonadj_med })
)
```

```{r adjacency_lorru}
adjacency_lorru_results <- analysis_f(
  function(df) with(df, {
    list(
      adj_med = median(df$lorr_du[df$adjacent]),
      nonadj_med = median(df$lorr_du[!df$adjacent])
    )
  }),
  function(model) with(model, { adj_med - nonadj_med }),
  function(model) with(model, { (adj_med - nonadj_med) / nonadj_med })
)
```

```{r adjacency_with_rlm}
adjacency_rlm_results <- analysis_f(
  function(df) rlm(dr_du ~ adjacent, data = df),
  function(model) coef(model)['adjacentTRUE'],
  function(model) coef(model) %>% { .['adjacentTRUE'] / .['(Intercept)'] }
)
```

```{r adjacency_with_covariates}
adjacency_covariates_results <- analysis_f(
  function(df) rlm(dr_du ~ adjacent + d_income + d_temperature + d_density, data = df),
  function(model) coef(model)['adjacentTRUE'],
  function(model) coef(model) %>% { .['adjacentTRUE'] / .['(Intercept)'] }
)
```

```{r commuting_analysis}
commuting_results <- analysis_f(
  function(df) rlm(dr_du ~ f_commuting, data = df),
  function(model) coef(model)['f_commuting'] * 1e-4,
  function(model) coef(model) %>% { .['f_commuting'] / .['(Intercept)'] * 1e-4 }
)
```

### Tables

```{r}
show_results <- function(df, caption) {
  df %>%
    mutate(
      coef_hci = 1.96 * coef_se * 0.5,
      coef_cil = coef - coef_hci,
      coef_ciu = coef + coef_hci,
      coef_star = coef_cil > 0 | coef_ciu < 0,
      ratio_hci = 1.96 * ratio_se * 0.5,
      ratio_cil = ratio - ratio_hci,
      ratio_ciu = ratio + ratio_hci,
      ratio_star = ratio_cil > 0 | ratio_ciu < 0
    ) %>%
    mutate_if(is.numeric, sigfig) %>%
    mutate_at(vars(ends_with('star')), ~ recode(as.numeric(.), `1` = '*', `0` = '')) %>%
    mutate(
      coef_display = str_glue('{coef} ({coef_cil} to {coef_ciu}){coef_star}'),
      ratio_display = str_glue('{ratio} ({ratio_cil} to {ratio_ciu}){ratio_star}')
    ) %>%
    select(dataset, coef_display, ratio_display) %>%
    kable(caption = caption)
}

show_results(
  adjacency_results,
  'Comparing adjacent and non-adjacent pairs. "Coef" is difference in median Δρ/Δτ between the two groups. "Ratio" is that difference divided by overall median.'
)

show_results(
  adjacency_lorru_results,
  'As above, but using LOR(ρ)/Δτ'
)

show_results(
  adjacency_rlm_results,
  'Using robust regression: Δρ/Δτ ~ Α. "Coef" is βΑ, "ratio" is βΑ/μ.'
)

show_results(
  commuting_results,
  'Robust regression on Δρ/Δτ ~ C, where C is the commuting fraction. "Coef" is β * 10^-4. "Ratio" is β * 10^-4 / μ.'
)
```

### Plots

```{r adjacency_plot}
boxplot_data_f <- function(df, ymin, ymax) {
  df %>%
    nest(-adjacent) %>%
    mutate(
      y = map(data, ~ .$dr_du),
      box = map(y, ~ boxplot.stats(.)$stats),
      boxplot_data = map(box, ~ tibble(
        ymin = max(.[1], ymin),
        lower = .[2],
        middle = .[3],
        upper = .[4],
        ymax = min(.[5], ymax)
      ))
    ) %>%
    select(adjacent, boxplot_data) %>%
    unnest()
}

boxplot_f <- function(cross_data, f_to_keep) {
  half_drop <- (1 - f_to_keep) / 2

  plot_data <- cross_data %>%
    mutate(
      y = map(cross_data, ~ .$dr_du),
      ymin = map_dbl(y, ~ quantile(., half_drop)),
      ymax = map_dbl(y, ~ quantile(., 1 - half_drop)),
      boxplot_data = pmap(list(cross_data, ymin, ymax), boxplot_data_f),
      point_data = pmap(list(cross_data, ymin, ymax), ~ filter(..1, between(dr_du, ..2, ..3)))
    )

  point_data <- plot_data %>%
    select(dataset, point_data) %>%
    unnest()

  boxplot_data <- plot_data %>%
    select(dataset, boxplot_data) %>%
    unnest()

  plot <- ggplot(data = NULL, aes(x = factor(adjacent))) +
    facet_wrap(~ dataset, scales = 'free_y') +
    geom_boxplot(
      data = boxplot_data,
      aes(lower = lower, upper = upper, middle = middle, ymin = ymin, ymax = ymax),
      stat = 'identity'
    ) +
    geom_jitter(data = point_data, aes(y = dr_du), size = 0.1, width = 0.2) +
    scale_x_discrete(
      '',
      labels = c(`TRUE` = 'Adjacent', `FALSE` = 'Not adj.')
    ) +
    ylab(expression(paste('Use-resistance association ', (Delta * rho / Delta * tau)))) +
    theme_cowplot() +
    theme(strip.background = element_blank())

  plot
}

adjacency_plot <- boxplot_f(cross_data, 0.90)
show(adjacency_plot)
ggsave('fig/adjacency_plot.pdf', plot = adjacency_plot)
ggsave('fig/adjacency_plot.png', plot = adjacency_plot)
```

```{r commuting_plot}
# swo: can include the rlm on the un-logged x values, as reported in the
# data, but they look weird and curvy when log-ing the x-vals

commute_plot <- cross_data %>%
  select(dataset, cross_data) %>%
  unnest() %>%
  group_by(dataset) %>%
  mutate(x = case_when(
    f_commuting == 0 ~ -6,
    TRUE ~ log10(f_commuting)
  )) %>%
  filter(between(ecdf(dr_du)(dr_du), 0.025, 0.975)) %>%
  ungroup() %>%
  ggplot(aes(x)) +
  facet_wrap(~ dataset, scales = 'free') +
  geom_point(aes(y = dr_du), shape = 1, size = 0.5) +
  #geom_line(aes(y = y)) +
  xlab('Commuting fraction (log10)') +
  ylab(expression(paste('Use-resistance association ', (Delta * rho / Delta * tau)))) +
  theme_cowplot() +
  theme(strip.background = element_blank())

show(commute_plot)
ggsave('fig/commute_plot.pdf', plot = commute_plot)
ggsave('fig/commute_plot.png', plot = commute_plot)
```
