---
title: "Theory of use-resistance relationships over geographic scales"
author: "Scott Olesen"
---

```{r global_options, echo = FALSE}
knitr::opts_chunk$set(
  fig.width = 7, fig.height = 4, fig.path = 'fig/',
  dev = c('png', 'pdf'),
  echo = FALSE, warning = FALSE, message = FALSE,
  cache = TRUE, autodep = TRUE)

pdf.options(useDingbats = FALSE, useKerning = FALSE)

library(kableExtra)
options(knitr.table.format = 'pandoc')
```

# Within-host neutral (WHN) model

```{r whn_parameters}
whn_base_parms = list(
  beta = 4.0,
  u = 1.0,
  tau50 = 1.5 / 12
)

# swo: always using davies base parms, so just bake those into the cost function?
# tau50 is the tau at which resistance prevalence is 50%
cost_function = function(beta, u, tau50) {
  k = 1.0

  1/2*(3*beta*k*tau50 - (3*k - 2)*tau50^2 - (k - 2)*u^2 + (beta*k - 4*(k - 1)*tau50)*u -
    sqrt(beta^2*k^2*tau50^2 + (k^2 - 4*k + 4)*tau50^4 + (k^2 - 4*k + 4)*u^4 - 2*(beta*k^2 -
    2*beta*k)*tau50^3 - 2*(beta*k^2 - 2*beta*k + 8*(k - 1)*tau50)*u^3 + (beta^2*k^2 -
    2*(k^2 + 12*k - 12)*tau50^2 + 2*(beta*k^2 + 6*beta*k)*tau50)*u^2 - 2*(beta^2*k^2*tau50 +
    8*(k - 1)*tau50^3 - (beta*k^2 + 6*beta*k)*tau50^2)*u))/(beta*k*tau50 - k*tau50^2 - k*u^2 +
    (beta*k - 2*k*tau50)*u)
}

whn_base_parms$cost = with(
  whn_base_parms,
  cost_function(beta, u, tau50)
)

whn_base_parms %>%
  as_tibble() %>%
  kable(caption = 'WHN model parameters')
```

## WHN single-population model

```{r whn_sim}
# vector to data frame
whn_unpack = function(x) {
  matrix(x, ncol = 4) %>%
    as_tibble() %>%
    setNames(c('X', 'S', 'R', 'D'))
}

# data frame to vector
whn_pack = function(df) {
  stopifnot(names(df) == c('X', 'S', 'R', 'D'))
  unlist(df)
}

whn_ode_func = function(t, state_vector, parms) {
  state = whn_unpack(state_vector)

  with(c(state, parms), {
    n_pop = nrow(state)
    stopifnot(dim(betaij) == rep(n_pop, 2))

    N = X + S + R + D

    dS = (betaij %*% ((S + D) / N)) * X - (u + taui) * S - (1 - cost) * (betaij %*% (R / N)) * S
    dR = (1 - cost) * (betaij %*% (R / N)) * X - u * R + taui * D
    dD = (1 - cost) * (betaij %*% (R / N)) * S - (u + taui) * D
    dX = -(dS + dR + dD)

    list(c(dX, dS, dR, dD))
  })
}

whn_sim = function(taui, epsilon) {
  n_pop = length(taui)

  parms = with(whn_base_parms, {
    betaij = matrix(beta * epsilon / n_pop, ncol = n_pop, nrow = n_pop)
    diag(betaij) <- beta * (1 - epsilon * (1 - 1 / n_pop))

    stopifnot(all(rowSums(betaij) == beta))
    stopifnot(all(colSums(betaij) == beta))

    c(whn_base_parms, list(
      taui = taui, epsilon = epsilon,
      betaij = betaij
    ))
  })

  state = data_frame(
    X = rep(0.990 / n_pop, n_pop),
    S = rep(0.005 / n_pop, n_pop),
    R = S,
    D = 0
  )

  state_vector = whn_pack(state)

  result = rootSolve::runsteady(state_vector, func = whn_ode_func, parms = parms)

  result$y %>%
    whn_unpack() %>%
    mutate(tau = taui,
           pop = seq_along(taui),
           rho = R / (S + R + D))
}
```

First, I reproduce Figure 2g with $k=1$.

```{r davies_repro}
whn_repro = whn_sim(taui = seq(0, 3 / 12, length.out = 16), epsilon = 0.0)

whn_repro %>%
  ggplot(aes(tau, rho)) +
  geom_point() +
  geom_line() +
  scale_x_continuous(name = 'antibiotic tx per month',
                     sec.axis = sec_axis(~ . * 12)) +
  ylab('proportion resistant')
```

```{r whn_single, eval = FALSE}
whn1 = crossing(
  n_pop = c(4, 8, 16),
  epsilon = c(0.00, 0.01, 0.10, 0.25, 0.50)
) %>%
  mutate(taui = map(n_pop, ~ seq(0.05, 0.2, length.out = .)),
         results = map2(taui, epsilon, whn_sim))
```

```{r whn_multi_plot, eval = FALSE}
whn_multipop_plot = whn1 %>%
  unnest() %>%
  ggplot(aes(taui, rho, color = factor(epsilon))) +
  facet_grid(~ n_pop) +
  geom_point() +
  geom_line()

whn_multipop_plot

ggsave('fig/whn_multipop.png', plot = whn_multipop_plot)
```

## WHN two-population model

```{r whn_double}
twopop_taui_df = data_frame(
  size = c('small', 'large'),
  taui = list(c(0.10, 0.15), c(0.05, 0.20))
)

whn2 = twopop_taui_df %>%
  crossing(epsilon = c(0, 0.01, 0.1, 0.25, 0.5, 1.0)) %>%
  mutate(results = map2(taui, epsilon, whn_sim))
```

```{r whn_double_results}
odds = function(p) p / (1 - p)
odds_ratio = function(p, q) odds(p) / odds(q)

whn2 %>%
  filter(epsilon > 0) %>%
  unnest() %>%
  mutate(rho = R / (S + R + D)) %>%
  group_by(epsilon, size) %>%
  mutate(delta_u = max(tau) - min(tau)) %>%
  group_by(epsilon, delta_u) %>%
  summarize(arr = max(rho) - min(rho),
            lor = log(odds_ratio(min(rho), max(rho)))) %>%
  ungroup() %>%
  mutate(arru = arr / delta_u,
         loru = lor / delta_u) %>%
  mutate_if(is.numeric, ~ round(., 2)) %>%
  arrange(epsilon, delta_u) %>%
  kable(caption = '2-population effect sizes, WHN model')

whn2 %>%
  filter(epsilon > 0) %>%
  unnest() %>%
  # swo: check this sum
  mutate(rho = R / (S + R + D)) %>%
  group_by(epsilon, size) %>%
  summarize_at('rho', mean) %>%
  arrange(epsilon, rev(size)) %>%
  mutate(rho = round(rho * 100, 1)) %>%
  kable(caption = 'Overall resistance levels')
```

## WHN nested populations

```{r whn_nest_sim}
whn_nest_sim = function(taui, eps_super, eps_sub) {
  n_pop = length(taui)
  pop_per = sqrt(n_pop)

  if (pop_per != as.integer(pop_per)) {
    stop('length of taui must be a perfect square')
  }

  superpop = rep(1:pop_per, each = pop_per)
  stopifnot(length(superpop) == length(taui))

  contact = outer(1:n_pop, 1:n_pop, Vectorize(function(i, j) {
    if (i == j) {
      (1 - eps_super * (1 - 1 / pop_per)) * (1 - eps_sub * (1 - 1 / pop_per))
    } else if (superpop[i] == superpop[j]) {
      (1 - eps_super * (1 - 1 / pop_per)) * eps_sub / pop_per
    } else {
      eps_super / pop_per * 1 / pop_per
    }
  }))

  stopifnot(all.equal(rowSums(contact), rep(1.0, n_pop)))
  stopifnot(all.equal(colSums(contact), rep(1.0, n_pop)))

  parms = with(whn_base_parms, {
    betaij = beta * contact

    c(whn_base_parms, list(
      taui = taui, eps_super = eps_super, eps_sub = eps_sub,
      betaij = betaij
    ))
  })

  state = data_frame(
    X = rep(0.990 / n_pop, n_pop),
    S = rep(0.005 / n_pop, n_pop),
    R = S,
    D = 0
  )

  state_vector = whn_pack(state)

  result = rootSolve::runsteady(state_vector, func = whn_ode_func, parms = parms)

  result$y %>%
    whn_unpack() %>%
    mutate(pop = seq_along(taui),
           superpop = superpop[pop],
           tau = taui,
           rho = R / (S + R + D))
}
```

```{r whn_nest_results}
whn_nest_taui = rep(seq(0.07, 0.18, length.out = 4), each = 4) + 0.02 * rep(seq(-1, 1, length.out = 4), 4)

whn_nest_results = crossing(
  eps_super = c(0.0, 0.01, 0.1, 0.5),
  eps_sub = c(0.0, 0.01, 0.1, 0.5)
) %>%
  filter(eps_super <= eps_sub) %>%
  mutate(data = map2(eps_super, eps_sub, ~ whn_nest_sim(whn_nest_taui, .x, .y)))
```

```{r whn_nest_display}
whn_nest_plot = whn_nest_results %>%
  unnest() %>%
  mutate_at(vars(starts_with('eps')), ~ sprintf('%.2f', .)) %>%
  ggplot(aes(tau * 12, rho)) +
  facet_grid(eps_sub ~ eps_super) +
  geom_hline(yintercept = 0.5, color = 'gray', linetype = 2) +
  geom_smooth(aes(color = NULL), method = 'lm', se = FALSE, color = 'black', size = 0.5) +
  geom_point(aes(shape = factor(superpop))) +
  scale_shape_manual(values = c(16, 1, 15, 0)) +
  scale_x_continuous(
    name = 'antibiotic use (yearly treatments per person)',
    limits = c(0.5, 2.5), breaks = c(0.5, 1.5, 2.5),
    sec.axis = sec_axis(
      trans = ~ .,
      name = expression(epsilon[super]),
      breaks = NULL
    )
  ) +
  scale_y_continuous(
    name = 'resistance (proportion of hosts)',
    limits = c(0.0, 1.01), expand = c(0, 0), breaks = c(0, 0.5, 1.0),
    sec.axis = sec_axis(
      trans = ~ .,
      name = expression(epsilon[sub]),
      breaks = NULL
    )
  ) +
  theme_classic() +
  theme(
    axis.title = element_text(size = 8),
    axis.text = element_text(size = 8, color = 'black'),
    panel.spacing = unit(3, 'mm'),
    legend.position = 'none',
    strip.background = element_blank(),
    strip.text = element_text(size = 8, color = 'black', face = 'bold'),
    # epsilon_arrow labels
    axis.title.x.top = element_text(size = 8, color = 'black'),
    axis.title.y.right = element_text(
      size = 8, color = 'black',
      # make right-hand label upright and up/down centered
      angle = 0, vjust = 0.5
    )
  )

show(whn_nest_plot)

ggsave('fig/whn_nest.pdf', plot = whn_nest_plot,
       width = 88, height = 110, units = 'mm')
```

```{r nested_model_uses}
data_frame(
  population = 1:16,
  superpopulation = rep(1:4, each = 4),
  tau = round(whn_nest_taui, 3)
) %>%
  kable(caption = 'Antibiotic use rates in nested models')
```

# D-types model

```{r dtypes_parameters}
dtypes_base_parms = list(
  n_d = 16,
  min_mu = 0.5,
  max_mu = 2.0,
  beta = 2.0,
  c_beta = 1.0,
  c_mu = 1.1,
  k = 15.0
)

dtypes_base_parms %>%
  as_tibble() %>%
  kable(caption = 'D-types base parameters')

dtypes_base_parms$mu_d = with(dtypes_base_parms,
                             seq(min_mu, max_mu, length = n_d))

dtypes_base_parms %$%
  data_frame(
    d = 1:n_d,
    mu_d = mu_d
  ) %>%
  kable(caption = 'D-types mu_d values')
```

## Single-population model

```{r dtype_helpers}
# pack the list into a vector
dtypes_pack = function(lst) {
  stopifnot(setequal(names(lst), c('X_i', 'S_id', 'R_id')))
  stopifnot(all(dim(lst$S_id) == dim(lst$R_id)))
  stopifnot(dim(lst$S_id)[1] == length(lst$X_i))
  c(lst$X_i, as.vector(lst$S_id), as.vector(lst$R_id))
}

# unpack the vector into a list
dtypes_unpack = function(x, n_pop, n_d) {
  stopifnot(length(x) == n_pop + 2 * (n_d * n_pop))

  x_end = n_pop
  s_start = x_end + 1
  s_end = x_end + n_d * n_pop
  r_start = s_end + 1
  r_end = s_end + n_d * n_pop
  stopifnot(r_end == length(x))

  list(X_i = x[1:x_end],
       S_id = matrix(x[s_start:s_end], nrow = n_pop, ncol = n_d),
       R_id = matrix(x[r_start:r_end], nrow = n_pop, ncol = n_d))
}

# for initially packing the data frame into the list
dtypes_df_to_list = function(df) {
  # check we have the right column names
  stopifnot(setequal(names(df), c('population', 'phenotype', 'dtype', 'value')))
  # check we have the right phenotypes
  stopifnot(setequal(df$phenotype, c('S', 'R', 'X')))

  # extract the no. populations and D-types
  n_pop = length(unique(df$population))
  n_d = length(unique(df$dtype))

  # there should be S and R for each pop/D-type combo
  stopifnot(nrow(filter(df, phenotype == 'S')) == n_pop * n_d)
  stopifnot(nrow(filter(df, phenotype == 'R')) == n_pop * n_d)
  # but X for only each pop
  stopifnot(nrow(filter(df, phenotype == 'X')) == n_pop)

  list(X_i = df %>% filter(phenotype == 'X') %$% values,
       S_id = df %>% filter(phenotype == 'S') %$% value %>% matrix(nrow = n_pop, ncol = n_d),
       R_id = df %>% filter(phenotype == 'R') %$% value %>% matrix(nrow = n_pop, ncol = n_d))
}

# for finally unpacking the list into a data frame
dtypes_list_to_df = function(lst) {
  # check for names
  stopifnot(setequal(names(lst), c('X_i', 'S_id', 'R_id')))

  # get dimensions
  n_pop = length(lst$X_i)
  n_d = dim(lst$S_id)[2]

  # check shapes
  stopifnot(all(dim(lst$S_id) == c(n_pop, n_d)))
  stopifnot(all(dim(lst$R_id) == c(n_pop, n_d)))

  X_rows = data_frame(pop = 1:n_pop, phenotype = 'X', dtype = NA, value = lst$X_i)
  S_rows = crossing(dtype = 1:n_d, pop = 1:n_pop) %>%
    mutate(phenotype = 'S', value = as.vector(lst$S_id))
  R_rows = crossing(dtype = 1:n_d, pop = 1:n_pop) %>%
    mutate(phenotype = 'R', value = as.vector(lst$R_id))

  bind_rows(X_rows, S_rows, R_rows) %>%
    arrange(pop, phenotype, dtype)
}
```

```{r dtypes_sim}
dtypes_ode_func = function(t, state_vector, parms) {
  state = dtypes_unpack(state_vector, n_pop = parms$n_pop, n_d = parms$n_d)

  with(c(state, parms), {
    stopifnot(length(X_i) == n_pop)
    stopifnot(dim(S_id) == c(n_pop, n_d))
    stopifnot(dim(R_id) == c(n_pop, n_d))

    # rowSums are over D-types (second index)
    v_id = (1.0 - ((S_id + R_id) / rowSums(S_id + R_id) - 1.0 / n_d)) ** k
    stopifnot(dim(v_id) == c(n_pop, n_d))

    N_i = X_i + rowSums(S_id + R_id)
    stopifnot(length(N_i) == n_pop)

    # "tau_i * S_id" does sum_i { tau_i S_id }, which is length P vector
    # to multiply by rows, need to do some fancy footwork: "mat %*% diag(row)"
    dS_id = v_id * (beta_ij %*% (S_id / N_i)) * X_i - tau_i * S_id - S_id %*% diag(mu_d)
    dR_id = v_id * (beta_ij %*% (R_id / N_i)) * X_i - c_mu * R_id %*% diag(mu_d)
    dX_i = -rowSums(dS_id + dR_id)

    list(dtypes_pack(list(X_i = dX_i, S_id = dS_id, R_id = dR_id)))
  })
}

dtypes_sim = function(tau_i, epsilon = 0) {
  n_pop = length(tau_i)

  parms = with(dtypes_base_parms, {
    beta_ij = matrix(beta * epsilon / n_pop, ncol = n_pop, nrow = n_pop)
    diag(beta_ij) <- beta * (1 - epsilon * (1 - 1 / n_pop))

    stopifnot(all(rowSums(beta_ij) == beta))
    stopifnot(all(colSums(beta_ij) == beta))

    c(dtypes_base_parms, list(
      n_pop = n_pop, tau_i = tau_i, epsilon = epsilon, beta_ij = beta_ij
    ))
  })

  n_d = parms$n_d

  state = list(X_i = rep(0.9 / n_pop, n_pop),
               S_id = matrix(0.05 / (n_pop * n_d), nrow = n_pop, ncol = n_d),
               R_id = matrix(0.05 / (n_pop * n_d), nrow = n_pop, ncol = n_d))

  state_vector = dtypes_pack(state)
  stopifnot(all.equal(sum(state_vector), 1))

  result = rootSolve::runsteady(state_vector, func = dtypes_ode_func, parms = parms,
                                stol = 1e-8 / n_pop, rtol = 1e-6 / n_pop, atol = 1e-6 / n_pop)

  result$y %>%
    dtypes_unpack(n_pop = n_pop, n_d = n_d) %>%
    dtypes_list_to_df() %>%
    left_join(data_frame(pop = 1:n_pop, tau = tau_i), by = 'pop')
}
```

First, I reproduce Figure 3 with $\tau = 0.075$.

```{r dtypes_repro_bars}
dtype_repro_bars = dtypes_sim(0.075, 0)

dtype_repro_bars %>%
  mutate(mu_d = dtypes_base_parms$mu_d[dtype]) %>%
  ggplot(aes(mu_d, value, fill = phenotype)) +
  geom_col() +
  xlab('Clearance rate') +
  ylab('Strain prevalence')
```

Next, I show the jagged line in Figure 4. It may not look exactly right because I ran a limited number of simulations:

```{r dtypes_repro_line}
dtype_repro_line = dtypes_sim(seq(0.04, 0.20, length.out = 50), 0)

dtype_repro_line %>%
  group_by(tau) %>%
  summarize(f_res = sum(value[phenotype == 'R']),
            f_all = sum(value)) %>%
  mutate(rho = f_res / f_all) %>%
  ggplot(aes(tau, rho)) +
  geom_point() +
  geom_line()
```

## D-types two-population

```{r dtypes_double}
dtypes2 = twopop_taui_df %>%
  crossing(epsilon = c(0, 0.01, 0.1, 0.25, 0.5, 1.0)) %>%
  mutate(results = map2(taui, epsilon, dtypes_sim))
```

```{r dtypes_double_results}
dtypes_simplify_results = function(df) {
  df %>%
    group_by(pop, phenotype) %>%
    summarize(
      value = sum(value),
      tau = unique(tau)
    ) %>%
    ungroup() %>%
    spread(phenotype, value) %>%
    mutate(rho = R / (R + S))
}

dtypes2 %>%
  mutate(results = map(results, dtypes_simplify_results)) %>%
  unnest() %>%
  filter(epsilon %in% c(0.01, 0.10, 0.25, 0.50)) %>%
  group_by(epsilon, size) %>%
  mutate(delta_u = max(tau) - min(tau)) %>%
  group_by(delta_u, epsilon) %>%
  summarize(
    lor = log(odds_ratio(min(rho), max(rho))),
    pct_res = mean(rho * 100)
  ) %>%
  ungroup() %>%
  mutate_if(is.numeric, ~ round(., 2)) %>%
  arrange(delta_u, epsilon) %>%
  kable(caption = 'D-types 2-population effect sizes and overall resistance levels')
```

## D-types nested

```{r dtypes_nest_sim}
dtypes_nest_sim = function(tau_i, eps_super, eps_sub) {
  n_pop = length(tau_i)
  pop_per = sqrt(n_pop)

  # swo: a lot of this code is reproduced, find a way to refactor
  if (pop_per != as.integer(pop_per)) {
    stop('length of taui must be a perfect square')
  }

  superpop = rep(1:pop_per, each = pop_per)
  stopifnot(length(superpop) == length(tau_i))

  contact = outer(1:n_pop, 1:n_pop, Vectorize(function(i, j) {
    if (i == j) {
      (1 - eps_super * (1 - 1 / pop_per)) * (1 - eps_sub * (1 - 1 / pop_per))
    } else if (superpop[i] == superpop[j]) {
      (1 - eps_super * (1 - 1 / pop_per)) * eps_sub / pop_per
    } else {
      eps_super / pop_per * 1 / pop_per
    }
  }))

  stopifnot(all.equal(rowSums(contact), rep(1.0, n_pop)))
  stopifnot(all.equal(colSums(contact), rep(1.0, n_pop)))

  parms = with(dtypes_base_parms, {
    c(dtypes_base_parms, list(
      beta_ij = beta * contact,
      n_pop = n_pop, tau_i = tau_i,
      eps_super = eps_super, eps_sub = eps_sub
    ))
  })

  n_d = parms$n_d

  state = list(X_i = rep(0.9 / n_pop, n_pop),
               S_id = matrix(0.05 / (n_pop * n_d), nrow = n_pop, ncol = n_d),
               R_id = matrix(0.05 / (n_pop * n_d), nrow = n_pop, ncol = n_d))

  state_vector = dtypes_pack(state)
  stopifnot(all.equal(sum(state_vector), 1))

  result = rootSolve::runsteady(state_vector, func = dtypes_ode_func, parms = parms,
                                stol = 1e-8 / n_pop, rtol = 1e-6 / n_pop, atol = 1e-6 / n_pop)

  result$y %>%
    dtypes_unpack(n_pop = n_pop, n_d = n_d) %>%
    dtypes_list_to_df() %>%
    left_join(data_frame(pop = 1:n_pop, superpop = superpop[pop], tau = tau_i), by = 'pop')
}
```

```{r dtypes_nest_results}
dtypes_nest_taui = rep(seq(0.07, 0.18, length.out = 4), each = 4) + 0.02 * rep(seq(-1, 1, length.out = 4), 4)

dtypes_nest_results = crossing(
  eps_super = c(0.0, 0.01, 0.1, 0.5),
  eps_sub = c(0.0, 0.01, 0.1, 0.5)
) %>%
  filter(eps_super <= eps_sub) %>%
  mutate(data = map2(eps_super, eps_sub, ~ dtypes_nest_sim(dtypes_nest_taui, .x, .y)))
```

```{r dtypes_nest_plot}
janky_superpop_f = function(pop) {
  n_superpop = sqrt(length(pop))
  rep(1:n_superpop, each = n_superpop)
}

dtypes_nest_plot = dtypes_nest_results %>%
  unnest() %>%
  group_by(eps_super, eps_sub, superpop, pop, phenotype) %>%
  summarize(
    value = sum(value),
    tau = unique(tau)
  ) %>%
  ungroup() %>%
  spread(phenotype, value) %>%
  mutate(rho = R / (S + R)) %>%
  ggplot(aes(tau * 12, rho)) +
  facet_grid(eps_sub ~ eps_super) +
  geom_hline(yintercept = 0.5, color = 'gray', linetype = 2) +
  geom_smooth(aes(color = NULL), method = 'lm', se = FALSE, color = 'black', size = 0.5) +
  geom_point(aes(shape = factor(superpop))) +
  scale_shape_manual(values = c(16, 1, 15, 0)) +
  scale_x_continuous(
    name = 'antibiotic use (yearly treatments per person)',
    limits = c(0.5, 2.5), breaks = c(0.5, 1.5, 2.5),
    sec.axis = sec_axis(
      trans = ~ .,
      name = expression(epsilon[super]),
      breaks = NULL
    )
  ) +
  scale_y_continuous(
    name = 'resistance (proportion of hosts)',
    limits = c(0.0, 1.01), expand = c(0, 0.1), breaks = c(0, 0.5, 1.0),
    sec.axis = sec_axis(
      trans = ~ .,
      name = expression(epsilon[sub]),
      breaks = NULL
    )
  ) +
  theme_classic() +
  theme(
    axis.title = element_text(size = 8),
    axis.text = element_text(size = 8, color = 'black'),
    panel.spacing = unit(3, 'mm'),
    legend.position = 'none',
    strip.background = element_blank(),
    strip.text = element_text(size = 8, color = 'black', face = 'bold'),
    # epsilon_arrow labels
    axis.title.x.top = element_text(size = 8, color = 'black'),
    axis.title.y.right = element_text(
      size = 8, color = 'black',
      # make right-hand label upright and up/down centered
      angle = 0, vjust = 0.5
    )
  )

show(dtypes_nest_plot)

ggsave('fig/dtypes_nest.pdf', plot = dtypes_nest_plot,
       width = 88, height = 110, units = 'mm')
```
