---
title: "Theory of use-resistance relationships over geographic scales"
author: "Scott Olesen"
output:
  pdf_document: default
header-includes:
    - \usepackage{bm}
    - \usepackage{mathrsfs}
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(
  fig.width=7, fig.height=4, fig.path='fig/',
  dev=c('pdf', 'png'),
  echo=FALSE, warning=FALSE, message=FALSE,
  cache=TRUE, autodep=TRUE)
pdf.options(useDingbats=FALSE, useKerning=FALSE)

library(kableExtra)
```

# Within-host neutral (WHN) model

```{r whn_parameters}
whn_base_parms = list(
  beta = 4.0,
  u = 1.0,
  tau50 = 1.5 / 12
)

# swo: always using davies base parms, so just bake those into the cost function?
# tau50 is the tau at which resistance prevalence is 50%
cost_function = function(beta, u, tau50) {
  k = 1.0
  
  1/2*(3*beta*k*tau50 - (3*k - 2)*tau50^2 - (k - 2)*u^2 + (beta*k - 4*(k - 1)*tau50)*u - 
    sqrt(beta^2*k^2*tau50^2 + (k^2 - 4*k + 4)*tau50^4 + (k^2 - 4*k + 4)*u^4 - 2*(beta*k^2 -
    2*beta*k)*tau50^3 - 2*(beta*k^2 - 2*beta*k + 8*(k - 1)*tau50)*u^3 + (beta^2*k^2 -
    2*(k^2 + 12*k - 12)*tau50^2 + 2*(beta*k^2 + 6*beta*k)*tau50)*u^2 - 2*(beta^2*k^2*tau50 +
    8*(k - 1)*tau50^3 - (beta*k^2 + 6*beta*k)*tau50^2)*u))/(beta*k*tau50 - k*tau50^2 - k*u^2 +
    (beta*k - 2*k*tau50)*u)
}

whn_base_parms$cost = with(
  whn_base_parms,
  cost_function(beta, u, tau50)
)

whn_base_parms %>%
  as_tibble() %>%
  kable(caption = 'WHN model parameters')
```

## WHN single-population model

**swo** reproduce the original Davies figure here

```{r whn_sim}
# vector to data frame
whn_unpack = function(x) {
  matrix(x, ncol = 4) %>%
    as_tibble() %>%
    setNames(c('X', 'S', 'R', 'D'))
}

# data frame to vector
whn_pack = function(df) {
  stopifnot(names(df) == c('X', 'S', 'R', 'D'))
  unlist(df)
}

whn_ode_func = function(t, state_vector, parms) {
  state = whn_unpack(state_vector)
  
  with(c(state, parms), {
    n_pop = nrow(state)
    stopifnot(dim(betaij) == rep(n_pop, 2))
    
    N = X + S + R + D
    
    dS = (betaij %*% ((S + D) / N)) * X - (u + taui) * S - (1 - cost) * (betaij %*% (R / N)) * S
    dR = (1 - cost) * (betaij %*% (R / N)) * X - u * R + taui * D
    dD = (1 - cost) * (betaij %*% (R / N)) * S - (u + taui) * D
    dX = -(dS + dR + dD)
    
    list(c(dX, dS, dR, dD))
  })
}

whn_sim = function(taui, epsilon) {
  n_pop = length(taui)

  parms = with(whn_base_parms, {
    betaij = matrix(beta * epsilon / n_pop, ncol = n_pop, nrow = n_pop)
    diag(betaij) <- beta * (1 - epsilon * (1 - 1 / n_pop))
    
    stopifnot(all(rowSums(betaij) == beta))
    stopifnot(all(colSums(betaij) == beta))
    
    c(whn_base_parms, list(
      taui = taui, epsilon = epsilon,
      betaij = betaij
    ))
  })
  
  state = data_frame(
    X = rep(0.990 / n_pop, n_pop),
    S = rep(0.005 / n_pop, n_pop),
    R = S,
    D = 0
  )
  
  state_vector = whn_pack(state)
  
  result = rootSolve::runsteady(state_vector, func = whn_ode_func, parms = parms)
  
  result$y %>%
    whn_unpack() %>%
    mutate(tau = taui,
           pop = seq_along(taui),
           rho = R / (S + R + D))
}
```

First, I reproduce Figure 2g with $k=1$.

```{r davies_repro}
whn_repro = whn_sim(taui = seq(0, 3 / 12, length.out = 16), epsilon = 0.0)

whn_repro %>%
  ggplot(aes(tau, rho)) +
  geom_point() +
  geom_line() +
  scale_x_continuous(name = 'antibiotic tx per month',
                     sec.axis = sec_axis(~ . * 12)) +
  ylab('proportion resistant')
```

```{r whn_single}
whn1 = crossing(
  n_pop = c(4, 8, 16),
  epsilon = c(0, 0.01, 0.1, 0.25, 0.50, 1.0)
) %>%
  mutate(taui = map(n_pop, ~ seq(0.05, 0.2, length.out = .)),
         results = map2(taui, epsilon, whn_sim))
```

```{r whn_single_plot}
whn_multipop_plot = whn1 %>%
  unnest() %>%
  ggplot(aes(taui, rho, color = factor(epsilon))) +
  facet_grid(~ n_pop) +
  geom_point() +
  geom_line()

whn_multipop_plot

ggsave('fig/whn_multipop.png', plot = whn_multipop_plot)
```

## WHN two-population model

```{r whn_double}
# swo: this is a bad name
whn2 = crossing(
    epsilon = c(0, 0.01, 0.1, 0.25, 0.5, 1.0),
    size = c('big', 'small')
) %>%
  mutate(taui = lapply(size, function(x) if(x == 'small') c(0.10, 0.15) else c(0.05, 0.20))) %>%
   mutate(results = map2(taui, epsilon, whn_sim))
```

```{r whn_double_results}
OR = function(p, q) (p / (1 - p)) / (q / (1 - q))

whn2 %>%
  filter(epsilon > 0) %>%
  unnest() %>%
  mutate(rho = R / (S + R + D)) %>%
  group_by(epsilon, size) %>%
  mutate(delta_u = max(taui) - min(taui)) %>%
  group_by(epsilon, delta_u) %>%
  summarize(rr = max(rho) - min(rho),
            lor = log(OR(min(rho), max(rho)))) %>%
  mutate(rru = rr / delta_u,
         loru = lor / delta_u)
```

# D-types model

```{r dtypes_parameters}
dtypes_base_parms = list(
  n_d = 16,
  min_mu = 0.5,
  max_mu = 2.0,
  beta = 2.0,
  c_beta = 1.0,
  c_mu = 1.1,
  k = 15.0
)

dtypes_base_parms %>%
  as_tibble() %>%
  kable(caption = 'D-types base parameters')

dtypes_base_parms$mu_d = with(dtypes_base_parms,
                             seq(min_mu, max_mu, length = n_d))

dtypes_base_parms %$%
  data_frame(
    d = 1:n_d,
    mu_d = mu_d
  ) %>%
  kable(caption = 'D-types mu_d values')
```

## Single-population model

```{r dtype_helpers}
# pack the list into a vector
dtypes_pack = function(lst) {
  stopifnot(setequal(names(lst), c('X_i', 'S_id', 'R_id')))
  stopifnot(all(dim(lst$S_id) == dim(lst$R_id)))
  stopifnot(dim(lst$S_id)[1] == length(lst$X_i))
  c(lst$X_i, as.vector(lst$S_id), as.vector(lst$R_id))
}

# unpack the vector into a list
dtypes_unpack = function(x, n_pop, n_d) {
  stopifnot(length(x) == n_pop + 2 * (n_d * n_pop))

  x_end = n_pop
  s_start = x_end + 1
  s_end = x_end + n_d * n_pop
  r_start = s_end + 1
  r_end = s_end + n_d * n_pop
  stopifnot(r_end == length(x))

  list(X_i = x[1:x_end],
       S_id = matrix(x[s_start:s_end], nrow = n_pop, ncol = n_d),
       R_id = matrix(x[r_start:r_end], nrow = n_pop, ncol = n_d))
}

# for initially packing the data frame into the list
dtypes_df_to_list = function(df) {
  # check we have the right column names
  stopifnot(setequal(names(df), c('population', 'phenotype', 'dtype', 'value')))
  # check we have the right phenotypes
  stopifnot(setequal(df$phenotype, c('S', 'R', 'X')))

  # extract the no. populations and D-types
  n_pop = length(unique(df$population))
  n_d = length(unique(df$dtype))

  # there should be S and R for each pop/D-type combo
  stopifnot(nrow(filter(df, phenotype == 'S')) == n_pop * n_d)
  stopifnot(nrow(filter(df, phenotype == 'R')) == n_pop * n_d)
  # but X for only each pop
  stopifnot(nrow(filter(df, phenotype == 'X')) == n_pop)

  list(X_i = df %>% filter(phenotype == 'X') %$% values,
       S_id = df %>% filter(phenotype == 'S') %$% value %>% matrix(nrow = n_pop, ncol = n_d),
       R_id = df %>% filter(phenotype == 'R') %$% value %>% matrix(nrow = n_pop, ncol = n_d))
}

# for finally unpacking the list into a data frame
dtypes_list_to_df = function(lst) {
  # check for names
  stopifnot(setequal(names(lst), c('X_i', 'S_id', 'R_id')))

  # get dimensions
  n_pop = length(lst$X_i)
  n_d = dim(lst$S_id)[2]

  # check shapes
  stopifnot(all(dim(lst$S_id) == c(n_pop, n_d)))
  stopifnot(all(dim(lst$R_id) == c(n_pop, n_d)))

  X_rows = data_frame(pop = 1:n_pop, phenotype = 'X', dtype = NA, value = lst$X_i) 
  S_rows = crossing(dtype = 1:n_d, pop = 1:n_pop) %>%
    mutate(phenotype = 'S', value = as.vector(lst$S_id))
  R_rows = crossing(dtype = 1:n_d, pop = 1:n_pop) %>%
    mutate(phenotype = 'R', value = as.vector(lst$R_id))

  bind_rows(X_rows, S_rows, R_rows) %>%
    arrange(pop, phenotype, dtype)
}
```

```{r dtypes_sim}
dtypes_ode_func = function(t, state_vector, parms) {
  state = dtypes_unpack(state_vector, n_pop = parms$n_pop, n_d = parms$n_d)

  with(c(state, parms), {
    stopifnot(length(X_i) == n_pop)
    stopifnot(dim(S_id) == c(n_pop, n_d))
    stopifnot(dim(R_id) == c(n_pop, n_d))

    # rowSums are over D-types (second index)
    v_id = (1.0 - ((S_id + R_id) / rowSums(S_id + R_id) - 1.0 / n_d)) ** k
    stopifnot(dim(v_id) == c(n_pop, n_d))

    N_i = X_i + rowSums(S_id + R_id)
    stopifnot(length(N_i) == n_pop)

    # "tau_i * S_id" does sum_i { tau_i S_id }, which is length P vector
    # to multiply by rows, need to do some fancy footwork: "mat %*% diag(row)"
    dS_id = v_id * (beta_ij %*% (S_id / N_i)) * X_i - tau_i * S_id - S_id %*% diag(mu_d)
    dR_id = v_id * (beta_ij %*% (R_id / N_i)) * X_i - c_mu * R_id %*% diag(mu_d)
    dX_i = -rowSums(dS_id + dR_id)

    list(dtypes_pack(list(X_i = dX_i, S_id = dS_id, R_id = dR_id)))
  })
}

dtypes_sim = function(tau_i, epsilon = 0) {
  n_pop = length(tau_i)

  parms = with(dtypes_base_parms, {
    beta_ij = matrix(beta * epsilon / n_pop, ncol = n_pop, nrow = n_pop)
    diag(beta_ij) <- beta * (1 - epsilon * (1 - 1 / n_pop))

    stopifnot(all(rowSums(beta_ij) == beta))
    stopifnot(all(colSums(beta_ij) == beta))

    c(dtypes_base_parms, list(
      n_pop = n_pop, tau_i = tau_i, epsilon = epsilon, beta_ij = beta_ij
    ))
  })
  
  n_d = parms$n_d
  
  state = list(X_i = rep(0.9 / n_pop, n_pop),
               S_id = matrix(0.05 / (n_pop * n_d), nrow = n_pop, ncol = n_d),
               R_id = matrix(0.05 / (n_pop * n_d), nrow = n_pop, ncol = n_d))

  state_vector = dtypes_pack(state)
  stopifnot(all.equal(sum(state_vector), 1))

  result = rootSolve::runsteady(state_vector, func = dtypes_ode_func, parms = parms,
                                stol = 1e-8 / n_pop, rtol = 1e-6 / n_pop, atol = 1e-6 / n_pop)

  result$y %>%
    dtypes_unpack(n_pop = n_pop, n_d = n_d) %>%
    dtypes_list_to_df() %>%
    left_join(data_frame(pop = 1:n_pop, tau = tau_i), by = 'pop')
}
```

First, I reproduce Figure 3 with $\tau = 0.075$.

```{r dtypes_repro_bars}
dtype_repro_bars = dtypes_sim(0.075, 0)

dtype_repro_bars %>%
  mutate(mu_d = dtypes_base_parms$mu_d[dtype]) %>%
  ggplot(aes(mu_d, value, fill = phenotype)) +
  geom_col() +
  xlab('Clearance rate') +
  ylab('Strain prevalence')
```

Next, I show the jagged line in Figure 4. It may not look exactly right because I didn't run a ton of intermediate points:

```{r dtypes_repro_line}
dtype_repro_line = dtypes_sim(seq(0.04, 0.20, length.out = 25), 0)

dtype_repro_line %>%
  group_by(tau) %>%
  summarize(f_res = sum(value[phenotype == 'R']),
            f_all = sum(value)) %>%
  mutate(rho = f_res / f_all) %>%
  ggplot(aes(tau, rho)) +
  geom_point() +
  geom_line()
```

## D-types two-population

```{r dtypes_double}
dtypes2 = crossing(
    epsilon = c(0, 0.01, 0.1, 0.25, 0.5, 1.0),
    size = c('big', 'small')
) %>%
  mutate(taui = lapply(size, function(x) if(x == 'small') c(0.10, 0.15) else c(0.05, 0.20))) %>%
   mutate(results = map2(taui, epsilon, whn_sim))
```

```{r dtypes_double_results}
dtypes2 %>%
  filter(epsilon > 0) %>%
  unnest() %>%
  mutate(rho = R / (S + R + D)) %>%
  group_by(epsilon, size) %>%
  mutate(delta_u = max(taui) - min(taui)) %>%
  group_by(epsilon, delta_u) %>%
  summarize(rr = max(rho) - min(rho),
            lor = log(OR(min(rho), max(rho)))) %>%
  mutate(rru = rr / delta_u,
         loru = lor / delta_u)
```
