---
title: "Theory of use-resistance relationships over geographic scales"
author: "Scott Olesen"
output:
  pdf_document: default
header-includes:
    - \usepackage{bm}
    - \usepackage{mathrsfs}
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(
  fig.width=7, fig.height=4, fig.path='fig/',
  dev=c('pdf'),
  echo=FALSE, warning=FALSE, message=FALSE,
  cache=TRUE, autodep=TRUE)
pdf.options(useDingbats=FALSE, useKerning=FALSE)
```

# To add

- Rework "mixing" coefficient like Blanquart *et al*.'s $\epsilon \in [0, 1 - 1/n]$, where $\beta_{i \to j} = \beta (1-\epsilon)$ if $i = j$ and $\beta \epsilon / (n-1)$ otherwise.
- Why don't resistances in the time analysis go all the way up to 1?
- Here I talk about changing the geographical scale of analysis; I think it might be more common to think about the geographical scale of transmission. If the scale of transmission is longer than the scale of the geographical analysis, then you're doing the same thing I talk about here, subdividing an island. (Conversely, if transmission scales are shorter than geographical scales, you might see a relationship.)
- I show that slopes within each island are more shallow than the cross-island relationship. It's also true that if you take a sample of within-island values from different islands, you'll also get a more shallow slope, just maybe not quite as shallow. I put that as a gray line.
- I do the derivation of the subpopulation equations by assuming that the target group is small so that the island-wide $(S, R, D, X)$ is fixed. In fact, you can say that the target group, which is a fraction $f$ of the whole island, has use $\tau_1$ such that the remaining $1-f$ has use $\tau_0$ where $f \tau_1 + (1-f) \tau_0 = \tau$. In this case, you get the same equations as below. In other words, so long as the subpopulations have use such that the island-wide average use is $\tau$, then the cross-island relationship holds regardless of the size of the subgroups.

## Outline

- Theory
  - Interactions between populations dilute $r ~ u$. Too small a scale gives low slopes.
  - Slopes, correlations, and $\sigma_x^2$. We expect slopes to be similar, so $r^2 : \sigma_x \to [-1, 1]$.
- MAUP
  - ECDC, IMS/NHSN, MarketScan/RO
  - Is there some level of aggregation that tends to deliver stronger slopes?
- Smaller scale
  - Medicare/RO: going down to HRRs isn't bad. I'm not sure this is very convincing.
- Figures
  - Model explanation, islands & ring illustration
  - Tuning parameters and connectivity; how connectivity affects the slope
  - Zonal & random samples, maybe in 3 separate plots? Or in one huge plot with lots of insets

# Introduction

We expect that changing antibiotic use may have an effect on resistance. One complicating factor in this potential relationship is geographic scale. At what scale ---individual, community, state, country--- must an intervention take place? Relatedly, at what geographic scale must an ecological analysis be conducted?

If the intervention or analysis is performed at a length scale much shorter than the length scale of transmission, then the outcome will be due mostly to the exposure of individuals other than the ones under study. The effect will be "washed out". On the other hand, if the analysis is on a length scale musch shorter than the biologically-relevant scale, then areas with different exposures and outcomes will end up mixed together, limiting the study's ability to identify the relationship.

To develop insight about this problem, I present a schematic model of how the geographical scale of analysis can affect the apparent relationship between antibiotic use and resistance.

# Methods

## Summary of approaches

We use *theoretical* and *empirical* approaches.

First, we run simulations where the length scale of transmission is known, and the analysis is varied between units shorter and longer than that scale.

Second, we re-analyze empirical, ecological-level data about use and resistance and determine which length scales of transmission are compatible with those data.

## Intuition behind the simulation approach

Consider a set of islands. The inhabitants only have contact with other members of their island. The inhabitants of each island have some levels of average antibiotic use and antibiotic resistance. For simplicity, imagine that the use-resistance pairs $(\tau, \rho)$ fall on a line, subject to some noise.

Clearly the correct scale of analysis is the island. Using a larger scale of analysis ---say, aggregating nearby islands and using their cross-island average use and resistance--- will decrease statistical power, as there are fewer independent units. On the other hand, too small a scale ---say, different populations on the same island--- means that the units are not independent, and the use-resistance relationship will be "washed out" by the population-average value.

To see this, consider (*i*) a "lower-using-than-average" islander with use $\tau_\mathrm{lo}$ on an island with average use $\tau_\mathrm{hi} > \tau_\mathrm{lo}$ and (*ii*) a "low-and-average" islander with use $\tau_\mathrm{lo}$ on an island with average use $\tau_\mathrm{lo}$. The lower-than-average islander, being surrounded by individuals with a higher average probability of resistant carriage, is more likely to carry a resistant strain than the low-and-average islander. Conversely, a "higher-using-than-average" islander will benefit from the lower resistance probability of their neighbors.

This logic helps explain why a hospitalized person, whether they are receiving antibiotics or not, is more likely to carry a resistant organism than a person with the same antibiotic use who is outside the hospital.

We will use simulations to quantify the change in the measured use-resistance relationship when changing the length scale of the analysis.

## Simulations and models

Generally, we are interested in the relationship between antibiotic use $\tau$ and antibiotic resistance $\rho$. In most of these models, $\rho := R / (1-X)$, where $R$ is the proportion of the population carrying resistant bacteria, and $X$ is the proportion that are not carriers.

I will run simulations using three analytical frameworks:

1. Completely separate islands, each with different $\tau$ and $\rho$.
1. A single large island, with one $(\tau, \rho)$ pair, that is strongly connected to small islands, each with its own $\tau$.
1. A single island, in which individuals are tracked by their time since last treatment.

As sensitivity analyses, I will also consider two more spatial arrangements:

- The large and small island simulation, but with the relative size of the two islands being modulated
- A ring of islands, each with their own $\tau$, with tunable connectivity

I will use a dynamical modeling framework, the "within-host neutrality" (WHN) model presented by [Davies *et al.*](https://www.biorxiv.org/content/early/2018/01/03/217232) (doi:10.1101/217232), thats exhibit coexistence between susceptible and resistant organisms, with some relationship between $\tau$ and $\rho$. As a sensitivity analysis, I will repeat the analyses using the model presented by Lehtinen *et al*. (hereafter "D-types" model).

## Model specifications

### Cross-island, WHN model

In the within-host neutrality model, contact between sensitive-colonized $S$ and resistant-colonized $R$ hosts can lead the $S$ hosts to become dual-colonized $D$. These $D$ hosts act like $S$ hosts for purposes of infection, but if they are treated with antibiotics, then they become $R$.
$$
\begin{aligned}
\dot{S} &= \beta (S + D) X - (u + \tau) S - k \beta \underline{(1-c)} R S \\
\dot{R} &= \beta (1-c) R X - u R + \tau D \\
\dot{D} &= k \beta (1-c) R S - (u + \tau) D \\
1 &= X + S + R + D
\end{aligned}
$$
where $X$ is uncolonized, $u$ is the natural recovery rate, $\tau$ is the antibiotic treatment rate, $k$ is the efficiency of multiple colonization, and $c$ is the transmission fitness cost of resistance. (The underlined term is erroneously missing in the original manuscript.)

The parameter values are drawn from the manuscript: $u = 1 \text{ month}^{-1}$, $\beta = 4$ (I presume the units are also month$^{-1}$, $\tau \in [0, 4]$ courses per person per year (i.e., up to $1/3 \text{ month}^{-1}$, and $k \in \{0, 1/4, 1/2, 1\}$. In the manuscript, the cost $c$ was chosen such that $\rho = 1/2$ when $\tau = 1.5$ courses per person per year. To find these values, I used a symbolic computation program (SageMath) to directly solve the differential equations at equilibrium.

```{r whn_parameters}
davies_base_parms = list(
  beta = 4.0,
  u = 1.0
)

# swo: always using davies base parms, so just bake those into the cost function?
# tau50 is the tau at which resistance prevalence is 50%
cost_function = function(beta, u, k, tau50) {
  stopifnot(between(k, 0, 1))
  
  ifelse(k==0,
    tau50 / (tau50 + u),
    1/2*(3*beta*k*tau50 - (3*k - 2)*tau50^2 - (k - 2)*u^2 + (beta*k - 4*(k - 1)*tau50)*u - 
      sqrt(beta^2*k^2*tau50^2 + (k^2 - 4*k + 4)*tau50^4 + (k^2 - 4*k + 4)*u^4 - 2*(beta*k^2 -
      2*beta*k)*tau50^3 - 2*(beta*k^2 - 2*beta*k + 8*(k - 1)*tau50)*u^3 + (beta^2*k^2 -
      2*(k^2 + 12*k - 12)*tau50^2 + 2*(beta*k^2 + 6*beta*k)*tau50)*u^2 - 2*(beta^2*k^2*tau50 +
      8*(k - 1)*tau50^3 - (beta*k^2 + 6*beta*k)*tau50^2)*u))/(beta*k*tau50 - k*tau50^2 - k*u^2 +
      (beta*k - 2*k*tau50)*u)
  )
}

# display the cost values
data_frame(
  k = c(0, 0.25, 0.5, 1)
) %>%
  mutate(cost = cost_function(4, 1, k, 1.5/12)) %>%
  kable()
```

As per the manuscript's method, the model is initiated with $X = 0.998$, $S = R = 0.001$, and $D = 0$, then run to equilibrium. The model is interrogated by altering $\tau$ and examining the equilibrium resistance prevalence $R/(1-X)$.

### Cross-island, D-type models

In the D-types model, there $n$ "duration"-types, each with clearance rate $\mu_i$. Each type $i$ has sensitive and resistant strains, so the compartments are uncolonized $X$, sensitive-colonized $S_i$, and $R_i$. The D-types are subject to balancing selection, encoded by
$$
v_i = \left( 1 - \left[ \frac{S_i + R_i}{\sum_j (S_j + R_j)} - \frac{1}{n} \right]\right)^k,
$$
where $k \geq 1$ is the force of that selection. Thus:
$$
\begin{aligned}
\dot{S}_i &= v_i \beta S_i X - (\tau + \mu_i) S_i \\
\dot{R}_i &= v_i \frac{\beta}{c_\beta} R_i X - c_\mu \mu_i R_i \\
\end{aligned}
$$
where $\beta$ is the transmission coefficient, $\tau$ is the antibiotic treatment rate, $c_\beta \geq 1$ and $c_\mu \geq 1$ are the fitness costs of resistance.

The parameter values are drawn from the paper: $n = 16$, $\beta = 2$, $c_\beta = 1$, $c_\mu = 1.1$, $k = 15$, $\tau \in [0.04, 0.20]$, $\mu_x$ evenly spaced in $[0.5, 2]$.

```{r dtypes_parameters}
# swo: some parms are lists, some are vectors. pick one?
dtypes_base_parms = list(
  n = 16,
  min_mu = 0.5,
  max_mu = 2.0,
  beta = 2.0,
  cbeta = 1.0,
  cmu = 1.1,
  k = 15.0
)

dtypes_base_parms$mux = with(dtypes_base_parms,
                             seq(min_mu, max_mu, length=n))
```

I initiated the model with $X=0.9$ and the remaining 10% of the population divided evenly among the $S_i$ and $R_i$. Again, $\tau$ is altered and the equilibrium resistance prevalence $\sum_i R_i / \sum_i (S_i + R_i)$ is examined.

### Small-island, WHN model

The big island has an antibiotic use $T$ that has fixed the population composition $(X, S, R, D)$ on that island. We stipulate that $X + S + R + D = 1$. The small island has use $\tau$ and composition $(x, s, r, d)$, also summing to unity. Given fixed big-island parameters, how does affecting the small-island $\tau$ change the small-island composition?

We assume that the small islands are small enough that changes in their use $\tau$ do not affect the large island or one another.
$$
\begin{aligned}
\dot{s} &= \beta (S + D) x - (u + \tau) s - k \beta (1-c) R s \\
\dot{r} &= \beta (1-c) R x - u R + \tau d \\
\dot{d} &= k \beta (1-c) R s - (u + \tau) d \\
1 &= x + s + r + d \\
\end{aligned}
$$

In these analyses, the big-island $T$ is set; the big-island equations are solved for and fixed at the equilibrium values $X$, $S$, $R$, and $D$; and then $\tau$ is varied and these equations are solved for the equilibrium, small-island resistance prevalence $\rho = \mathscr{r} / (1-x)$.

### Small-island, D-types model

The force of infection and the balancing selection
are determined by the big-island $S_i$ and $R_i$. People move between the small-island compartments $x$, $s_i$, and $r_i$:
$$
\begin{aligned}
\dot{s}_i &= V_i \beta S_i x - (\tau + \mu_i) S_i \\
\dot{r}_i &= V_i \frac{\beta}{c_\beta} R_i x - c_\mu \mu_i r_i \\
\end{aligned}
$$
The $V_i$ is the balancing force term, which I capitalized to emphasize that it depends on the big-island values.

The small-island resistance prevalence will depend on all the D-types:
$$
\rho  = \frac{\sum_i r_i}{\sum_i (s_i + r_i)}.
$$

### Time-since-treatment, WHN model

Another way to model this within-island difference is to consider the time between individuals' treatments. This requires modifying the equations so that individuals can transfer between compartments at a single timestep (via colonization or recovery) or can move between time since treatment (via treatment or by "aging out"). I modify the within-host model to show this:
$$
\begin{aligned}
\dot{S}_t &= \beta (S_\bullet + P_\bullet) X_t - (u + \tau) S_t - k \beta (1-c) R_\bullet S_t - \mathbf{1}_{t \neq T} \alpha S_t + \mathbf{1}_{t \neq 1} \alpha S_{t-1} \\
\dot{R}_t &= \beta (1-c) R_\bullet X_t - (u + \tau) R_t + \mathbf{1}_{t=1} \tau (R_\bullet + D_\bullet) - \mathbf{1}_{t \neq T} \alpha R_t + \mathbf{1}_{t \neq 1} \alpha R_{t-1} \\
\dot{D}_t &= k \beta (1-c) R_\bullet S_t - (u + \tau) D_t - \mathbf{1}_{t \neq T} \alpha D_t + \mathbf{1}_{t \neq 1} \alpha D_{t-1} \\
\dot{X}_t &= -\beta (S_\bullet + D_\bullet) X_t - \beta (1-c) R_\bullet X_t + u (S_t + R_t + D_t) - \tau X_t + \\
  &\qquad \mathbf{1}_{t=1} \tau (S_\bullet + X_\bullet) - \mathbf{1}_{t \neq T} \alpha X_t + \mathbf{1}_{t \neq 1} \alpha X_{t-1}
\end{aligned}
$$
where the subscript $t$ indicates time since treatment ($t=1$ means 0 to 1 month since treatment, $t=2$ means 1-2 months, etc.), $T$ is the number of time brackets recorded, the bullet indicates summation ($S_\bullet = \sum_t S_t$), $\mathbf{1}$ is the indicator function, and $1/\alpha = 1 \text{ month}$ is the duration of the time bracket. Note that treatment moves individuals out of the $X_t$ and $R_t$ compartments not because they change colonization status but because they move individuals back to $X_1$ and $R_1$.

I set $T = 72$ (i.e., recording up to 6 years since treatment). These equations are also solved numerically, setting all initial populations into the "just treated" compartments $t=1$.

The resistance propotion $\rho_t = R_t / (S_t + R_t + D_t)$ depends on the time bracket. Individuals in bracket $t$, who were treated at most $t/\alpha$ ago, have "average use" $\alpha/t$.

# Results

## Cross-island, WHN model

I reproduce Figure 2e from Davies *et al*., showing that resistance prevalence increases with antibiotic use, and a greater efficiency of co-colonization $k$ allows greater range of coexistence.

```{r davies_cross_sim}
# within-host (Davies) cross-island simulation
davies_cross_sim = function(tau, k, tau50=1.5/12) {
  parms = with(davies_base_parms,
               c(davies_base_parms, tau=tau, k=k,
                 cost=cost_function(beta, u, k, tau50)))
  
  # initial state
  state = c(X = 0.998, S = 0.001, R = 0.001, D = 0.0)
  stopifnot(sum(state) == 1.0)
  
  ode_func = function(t, state, parms) {
    with(c(state, parms), {
      # NB: The (1-c) term in the last term of dS/dt
      dS = beta * (S + D) * X - (u + tau) * S - k * (1 - cost) * beta * R * S
      dR = beta * (1 - cost) * R * X - u * R + tau * D
      dD = k * beta * (1 - cost) * R * S - (u + tau) * D
      dX = -(dS + dR + dD)
      
      stopifnot(abs(dS + dR + dD + dX) < 1e-6)
      list(c(dX, dS, dR, dD))
    })
  }

  result = rootSolve::runsteady(state, func=ode_func, parms=parms)
  result$y %>% as.list %>% as_tibble
}

davies_cross_all = crossing(
  tau = round(seq(0, 4, length.out=51) / 12, 3),
  k = c(0, 0.25, 0.5, 1.0)
) %>%
  mutate(model = map2(tau, k, davies_cross_sim)) %>%
  unnest() %>%
  mutate(rho = R / (1 - X))

davies_cross = davies_cross_all %>% filter(k == 1.0)

davies_cross_all %>%
  ggplot(aes(tau, rho, color=factor(k), group=factor(k))) +
  geom_point(shape=0) +
  geom_line() +
  scale_x_continuous('Antibiotic courses per person per month (tau)',
                     sec.axis = sec_axis(name='Antibiotic courses per year', ~ . * 12)) +
  ylab('Resistance prevalence') +
  theme_minimal()
```

In the remainder of this paper, we will stick to $k=1$ and set $\tau_{50} = 1.5 / 12$.

## Small-island, WHN model

```{r davies_small_sim}
davies_small_sim = function(big_tau, tau, k = 1.0, tau50 = 1.5/12) {
  # get the big-island results
  big = davies_cross_sim(big_tau, k)
  
  parms = with(davies_base_parms,
               c(davies_base_parms, tau = tau,
                 k = k, cost = cost_function(beta, u, k, tau50)))

  # initial state is the big-island results
  state = unlist(big)
  stopifnot(abs(sum(state) - 1) < 1e-6)
    
  ode_func = function(t, state, parms) {
    with(c(state, parms), {
      dS = beta * (big$S + big$D) * X - (u + tau) * S - k * beta * (1 - cost) * big$R * S
      dR = beta * (1 - cost) * big$R * X - u * R + tau * D
      dD = k * beta * (1 - cost) * big$R * S - (u + tau) * D
      dX = -(dS + dR + dD)

      stopifnot(abs(sum(dX + dS + dR + dD)) < 1e-6)
      list(c(dX, dS, dR, dD))
    })
  }

  result = rootSolve::runsteady(state, func = ode_func, parms = parms)
  result$y %>% as.list %>% as_tibble
}

davies_small = crossing(big_tau = round(c(1.25, 1.50, 1.75) / 12, 3),
                        delta_tau = seq(-0.1, 0.1, length.out = 11)) %>%
  mutate(tau = big_tau + delta_tau) %>%
  mutate(model = map2(big_tau, tau, davies_small_sim)) %>%
  unnest() %>%
  mutate(rho = R / (1 - X))
```

```{r davies_small_plot}
davies_small %>%
  ggplot(aes(tau, rho)) +
  geom_line(data = davies_cross) +
  geom_vline(aes(xintercept = big_tau), color = 'gray', linetype = 2) +
  geom_point(aes(color = factor(big_tau))) +
  geom_line(aes(color = factor(big_tau)))
```

## Time-since-treatment, WHN model

```{r davies_time_sim}
drop_last = function(x) x[-length(x)]

davies_time_sim = function(tau, k = 1.0, tau50 = 1.5 / 12, alpha = 1.0, nT = 72) {
  # nT = number of time intervals
  parms = with(davies_base_parms,
             c(davies_base_parms, tau=tau,
               alpha = alpha, nT = nT,
               k = k, cost = cost_function(beta, u, k, tau50)))

  # state is a matrix:
  # rows are X, S, R, D (=S_R); columns are time brackets
  # set all compartments to 0 except for the "just-treated" bracket
  state = matrix(0, nrow = 4, ncol = parms$nT)
  state[,1] = c(0.998, 0.001, 0.001, 0.0)
  state_c = as.vector(state)
  stopifnot(sum(state_c) == 1.0)
  
ode_func = function(t, state_c, parms) {
    with(parms, {
      # unpack the state
      state = matrix(state_c, nrow = 4)
      stopifnot(ncol(state) == nT)
      X = state[1, ]
      S = state[2, ]
      R = state[3, ]
      D = state[4, ]
      
      Xo = sum(X)
      So = sum(S)
      Ro = sum(R)
      Do = sum(D)
      
      dX = -beta * (So + Do) * X - beta*(1-cost) * Ro * X + u * (S + R + D) - tau * X + c(tau * (So + Xo), rep(0, nT - 1)) - c(alpha * drop_last(X), 0) + c(0, alpha * drop_last(X))
      dS = beta * (So + Do) * X - (u + tau) * S - k * beta * (1-cost) * Ro * S - c(alpha * drop_last(S), 0) + c(0, alpha * drop_last(S))
      dR = beta*(1-cost) * Ro * X - (u + tau) * R + c(tau * (Ro + Do), rep(0, nT - 1)) - c(alpha * drop_last(R), 0) + c(0, alpha * drop_last(R))
      dD = k * beta*(1-cost) * Ro * S - (u + tau) * D - c(alpha * drop_last(D), 0) + c(0, alpha * drop_last(D))
      
      dstate = rbind(dX, dS, dR, dD)
      dstate_c = as.vector(dstate)
      stopifnot(abs(sum(dstate_c)) < 1e-6)
      list(dstate_c)
    })
  }

  result = rootSolve::runsteady(state_c, func = ode_func, parms = parms)
  end_state = result$y %>% matrix(nrow=4)
  
  time = 1:parms$nT
  bind_rows(
    data_frame(compartment = 'X', time, y = end_state[1,]),
    data_frame(compartment = 'S', time, y = end_state[2,]),
    data_frame(compartment = 'R', time, y = end_state[3,]),
    data_frame(compartment = 'D', time, y = end_state[4,])
  )
}

davies_time = crossing(tau = round(c(1.25, 1.5, 1.75) / 12, 3)) %>%
  mutate(model = map(tau, davies_time_sim)) %>%
  unnest() %>%
  rename(big_tau = tau) %>%
  spread(compartment, y) %>%
  mutate(tau = 1 / time,
         rho = R / (S + R + D))
```

```{r davies_time_plot}
davies_time %>%
  ggplot(aes(tau, rho)) +
  geom_line(data = davies_cross) +
  geom_point(aes(color = factor(big_tau)), shape = 2) +
  geom_line(aes(color = factor(big_tau)), linetype = 2)
```

For the time simulation, I verify that there are enough time compartments so that they long-time-since-treatment compartments are nearly empty. The columns are the $\tau$.

```{r confirm_time_comparments}
davies_time %>%
  gather('compartment', 'value', D:X) %>%
  ggplot(aes(time, value, color = compartment)) +
  geom_line() +
  facet_grid(~ big_tau) +
  xlab('Time since treatment (months)') +
  ylab('Proportion')
```

I compare the cross-island, within-island fixed-population, and within-island recorded-time simulations by plotting all the results on the same plot. The black lines show the cross-island results (i.e., the results from Davies *et al*. Figure 2e, above). The circles show results from the within-island fixed-population simulation. The colored lines show the within-island recorded-time simulation. The color of the circles and lines shows the island-level average $\tau$ used in the simulations. (You can read off those values by looking at where the colored lines intercept the black lines: when the island-level $\tau$ and within-island use are the same, you get the same results.) Panels show values of co-colonization efficiency $k$.

```{r}
davies_time_comp = davies_time %>% filter(tau < 0.3)

ggplot(data = NULL, aes(tau, rho)) +
  geom_line(data = filter(davies_cross, k == 1.0)) +
  geom_point(data = davies_small, aes(color = factor(big_tau)), shape = 1) +
  geom_line(data = davies_small, aes(color = factor(big_tau))) +
  geom_point(data = davies_time_comp, aes(color = factor(big_tau)), shape = 2) +
  geom_line(data = davies_time_comp, aes(color = factor(big_tau)), linetype = 2)
```

## Cross-island, D-types model

```{r dtype_cross}
dtypes_ode_func = function(t, Ixp, parms) {
  with(parms, {
    Ixs = Ixp[c(TRUE, FALSE)]
    Ixr = Ixp[c(FALSE, TRUE)]
    Ix = Ixs + Ixr
    U = 1.0 - sum(Ix)
    
    vx = (1.0 - (Ix / sum(Ix) - 1.0 / n)) ** k
    dIxs = beta * Ixs * vx * U - (tau + mux) * Ixs
    dIxr = beta / cbeta * Ixr * vx * U - cmu * mux * Ixr
    
    list(c(rbind(dIxs, dIxr)))
  })
}

dtypes_cross_sim = function(tau) {
  parms = c(dtypes_base_parms, tau=tau)
  y = rep(0.1 / 2 * parms$n, 2 * parms$n)
  result = rootSolve::runsteady(y, func=dtypes_ode_func, parms=parms)
  
  data_frame(dtype=rep(1:parms$n, rep(2, parms$n)),
             phenotype=rep(c('S', 'R'), parms$n),
             y=result$y)
}

dtypes_cross = data_frame(tau = seq(0.04, 0.20, length.out = 200)) %>%
  mutate(model = map(tau, dtypes_cross_sim)) %>%
  unnest() %>%
  group_by(tau, phenotype) %>%
  summarize(y = sum(y)) %>%
  ungroup() %>%
  spread(phenotype, y) %>%
  mutate(rho = R / (S + R))
```

I reproduce Figure 3 from Lehtinen *et al*.:

```{r dtypes_cross_plot}
dtypes_cross_sim(0.075) %>%
  ggplot(aes(factor(dtype), y, fill=phenotype)) +
  geom_col(position='dodge') +
  xlab('D-type') +
  ylab('proportion')
```

I overlay the cross-island (black) and within-island (i.e., fixed-population; colored) results. The black line reproduces Lehtinen *et al*. Figure 4 ( with $n=16$).

```{r dtypes_small}
dtypes_small_sim = function(island_tau, tau) {
  # compute the island-level results
  island_results = dtypes_cross_sim(island_tau)
  island_Ixp = island_results$y
  
  island_Ixs = island_Ixp[c(TRUE, FALSE)]
  island_Ixr = island_Ixp[c(FALSE, TRUE)]
  island_Ix = island_Ixs + island_Ixr
  island_vx = with(dtypes_base_parms,
                   (1.0 - (island_Ix / sum(island_Ix) - 1.0 / n)) ** k)
  
  ode_func = function(t, Ixp, parms) {
    with(parms, {
      Ixs = Ixp[c(TRUE, FALSE)]
      Ixr = Ixp[c(FALSE, TRUE)]
      Ix = Ixs + Ixr
      U = 1.0 - sum(Ix)
    
      dIxs = beta * island_Ixs * island_vx * U - (tau + mux) * Ixs
      dIxr = beta / cbeta * island_Ixr * island_vx * U - cmu * mux * Ixr
    
      list(c(rbind(dIxs, dIxr)))
    })
  }
  
  # swo: easier to read with parms= right next to result=
  parms = c(dtypes_base_parms, tau=tau)
  result = rootSolve::runsteady(island_Ixp, func=ode_func, parms=parms)
  
  data_frame(
    dtype=rep(1:parms$n, rep(2, parms$n)),
    phenotype=rep(c('S', 'R'), parms$n),
    y=result$y
  )
}

dtypes_small = crossing(
    big_tau = c(0.075, 0.125, 0.175),
    delta_tau = seq(-0.025, 0.025, length.out = 11)
) %>%
  mutate(tau = big_tau + delta_tau) %>%
  mutate(model = map2(big_tau, tau, dtypes_small_sim)) %>%
  unnest() %>%
  group_by(big_tau, tau, phenotype) %>%
  summarize(y = sum(y)) %>%
  ungroup() %>%
  spread(phenotype, y) %>%
  mutate(rho = R / (S + R))
```

```{r dtypes_small_plot}
dtypes_small %>%
  ggplot(aes(tau, rho)) +
  geom_line(data = dtypes_cross) +
  geom_point(aes(color = factor(big_tau))) +
  geom_line(aes(color = factor(big_tau)))
```

# Discussion

As expected the slopes are shallower within islands, but they are a *lot* shallower.

This is especially stark in the D-types model. In that model, coexistence comes from the fact that there are a range of clearance rates $\mu_x$. Some of these rates support sensitives and some support resistants. Whether an individual personally use more or less antibiotics doesn't flip the entire population of D-types. It appears that the individual-level variations in use only weakly modulate the force of infection an individual feels from the rest of the population.

In the within-host model, the within-island slopes are not so immensely shallow, but they are markedly more shallow than the cross-island relationship. The results of the two within-island simulations (fixed-population and time-since-treatment) give nearly identical results.

These results suggest that the best way to change your risk of antibiotic resistance is to change your environment, not your antibiotic use. I always thought that the worst thing about going to the hospital isn't that you're likely to use more antibiotics; it's that you're exposed to way more resistant bugs.

The fact that the within-island slopes are really shallow implies that, if an analysis did go below the relevant length scale, you'd expect a quick drop-off in the slopes.

This toy moel is limited because there's only one length scale. In reality there aren't completely-separated, well-mixed islands; instead there's a network of length scales. I expect that having multiple length scales would soften this weaking of slopes with shorter length scales so that you don't see a sudden cliff.

# Addenda

## Two length scales

Divide an island into islets. Each of the $N$ islets $i$ has use $\tau_i$ such that $\tau = (1/N)\sum_i \tau_i$ and size $\phi_i$ such that $\sum_i \phi_i = 1$. Introduce some parameter $\alpha \in [0, 1]$ that encodes the amount of mixing between the islets. For $\alpha = 0$, the islets are completely separate; for $\alpha = 1$, they are all completely mixed. Setting $\alpha = 1$ reproduces the cross-island results, while setting $\alpha = 0$ creates new cross-island-type results, only with the islet values $\tau_i$ rather than the island-wide values $\tau$.

$$
\begin{aligned}
\dot{S}_i &= \beta (S_i' + D_i') X_i - (u + \tau_i) S_i - k \beta (1-c) R_i' S_i \\
\dot{R}_i &= \beta R_i' X_i - u R_i + \tau_i D_i \\
\dot{D}_i &= k \beta (1-c) R_i' S_i - (u + \tau_i) D_i \\
\phi_i &= X_i + S_i + R_i + D_i \\
1 &= \sum_i \phi_i \\
\tau &= (1/N) \sum_i \tau_i
\end{aligned}
$$
where the primes indicate the scaled, cross-islet forces of infection:
$$
S_i' \equiv (1-\alpha) \frac{S_i}{\phi_i} + \alpha \sum_j S_j
$$
and similarly for $R_i'$ and $D_i'$.

```{r}
islet_v2l = function(x) {
  x %>%
    matrix(ncol=4, dimnames=list(NULL, c('X', 'S', 'R', 'D'))) %>%
    as_tibble() %>%
    as.list()
}

islet_ode_func = function(t, state_vector, parms) {
  state = islet_v2l(state_vector)
  
  with(c(state, parms), {
    prime_f = function(x) (1 - alpha) * x / phii + alpha * sum(x)
    
    Sprime = prime_f(S)
    Rprime = prime_f(R)
    Dprime = prime_f(D)
    
    dS = beta * (Sprime + Dprime) * X - (u + taui) * S - k * beta * (1 - cost) * Rprime * S
    dR = beta * (1 - cost) * Rprime * X - u * R + taui * D
    dD = k * beta * (1 - cost) * Rprime * S - (u + taui) * D
    dX = -(dS + dR + dD)
    
    change_vector = c(dX, dS, dR, dD)
    
    stopifnot(abs(sum(change_vector)) < 1e-12)

    list(c(dX, dS, dR, dD))
  })
}

islet_sim = function(taui, k, alpha, tau50=1.5/12) {
  n_islets = length(taui)
  phii = rep(1 / n_islets, n_islets)
  
  stopifnot(length(phii) == length(taui))
  stopifnot(sum(phii) == 1.0)
  
  parms = with(davies_base_parms,
               c(davies_base_parms, taui=list(taui), phii=list(phii), alpha=alpha,
                 k=k, cost=cost_function(beta, u, k, tau50)))
  
  state = data_frame(
    X = phii * 0.990,
    S = phii * 0.005,
    R = phii * 0.005,
    D = phii * 0
  )
  
  state_vector = unlist(state)
  
  result = rootSolve::runsteady(state_vector, func=islet_ode_func, parms=parms)
  
  result$y %>%
    islet_v2l() %>%
    as_tibble() %>%
    mutate(tau = taui, islet = seq_along(phii))

}

islets = crossing(big_tau = round(c(1.25, 1.50, 1.75) / 12, 3),
                  delta_tau = seq(-0.1, 0.1, length.out = 11),
                  k = c(1.0),
                  alpha = c(0.0, 0.1, 0.5, 1.0)) %>%
  mutate(tau = big_tau + delta_tau) %>%
  group_by(big_tau, alpha, k) %>%
  do(islet_sim(taui=.$tau, k=unique(.$k), alpha=unique(.$alpha))) %>%
  ungroup() %>%
  mutate(rho = R/(S + R + D))

# compare with within-host neutral model, within-island (subpop) model
# dotted lines
islets %>%
  ggplot(aes(tau, rho, color = factor(big_tau))) +
  geom_point() + geom_line() +
  geom_line(data = davies_cross, color = 'black') +
  geom_line(data = davies_small, linetype = 2) +
  facet_wrap(~ alpha)
```

Colored points and lines show the results from this simulation. Grey boxes show values of the mixing parameter $\alpha$. Black lines show the cross-island simulations. Dotted lines show the subpopulation simulations. I expect that the cross-island and cross-islet simulations coincide for $\alpha = 0$ and that the cross-islet and subpopulation models will be similar when $\alpha = 1$.

## Tunable geographical influence

As a variation on the previous, imagine that the islands can interact with one another in a tunable way. First, arrange the $N$ islands in a ring. The interaction between two islands $i$ and $j$ will be a function of the distance between them $d(i, j) = \min( |i-j|, N - |i-j| )$ as well as a separation parameter $\alpha$. For $\alpha \to \infty$, the islands are completely separate; for $\alpha = 0$, the islands interact fully. Terms like $\beta X S$ in the original model become like:
$$
\beta S_i \sum_j f_{ij} \frac{I_j}{\phi_j},
$$
where the $f_{ij}$ is the proportion of encounters made by people on island $i$ with people on island $j$. For exponential decay in interaction, this is:
$$
f_{ij} = \frac{ \exp\{ -\alpha \, d(i,j) \} }{\sum_{j=1}^N \exp \{ -\alpha \, d(i,j) \} }.
$$

For $\alpha = 0$, $f_{ij} = 1/N$ for all $i$ and $j$. For $\alpha \to \infty$, $W_{ij} = \mathbf{1}_{i=j}$.

It's easier to think about using a different tuning parameters, which is the fraction of contacts that are made on the island, which is $1 / \sum_j e^{-\alpha d}$, which uniquely determines $\alpha$.

Here I use 50 islands, with $\tau$ going linearly from 0 at point in the ring to 3 courses per month at the other end. Then I tune $\alpha$ to see how mixing dilutes the use-resistance relationship.

```{r}
ring_dist = function(i, j, n) pmin(abs(i - j), n - abs(i - j))

f_ij_f = function(n, alpha) {
  f_ij_denom = sum(exp(-alpha * ring_dist(1, 1:n, n)))
  outer(1:n, 1:n, function(i, j) {
    exp(-alpha * ring_dist(i, j, n))
  }) %>%
    divide_by(f_ij_denom)
}

ring_ode_func = function(t, state_vector, parms) {
  state = islet_v2l(state_vector)
  
  with(c(state, parms), {
    n = length(phii)
    
    f_ij = f_ij_f(n, alpha)
    prime_f = function(x) f_ij %*% (x / phii)
    
    Sprime = prime_f(S)
    Rprime = prime_f(R)
    Dprime = prime_f(D)
    
    dS = beta * (Sprime + Dprime) * X - (u + taui) * S - k * beta * (1 - cost) * Rprime * S
    dR = beta * (1 - cost) * Rprime * X - u * R + taui * D
    dD = k * beta * (1 - cost) * Rprime * S - (u + taui) * D
    dX = -(dS + dR + dD)
    
    change_vector = c(dX, dS, dR, dD)
    
    stopifnot(abs(sum(change_vector)) < 1e-12)

    list(c(dX, dS, dR, dD))
  })
}

ring_sim = function(taui, alpha, k = 1.0, tau50 = 1.5/12) {
  n_islets = length(taui)
  phii = rep(1 / n_islets, n_islets)
  
  stopifnot(length(phii) == length(taui))
  stopifnot(sum(phii) == 1.0)
  
  parms = with(davies_base_parms,
               c(davies_base_parms, taui=list(taui), phii=list(phii), alpha=alpha,
                 k=k, cost=cost_function(beta, u, k, tau50)))
  
  state = data_frame(
    X = phii * 0.990,
    S = phii * 0.005,
    R = phii * 0.005,
    D = phii * 0
  )
  
  state_vector = unlist(state)
  
  result = rootSolve::runsteady(state_vector, func=ring_ode_func, parms=parms)
  
  result$y %>%
    islet_v2l() %>%
    as_tibble() %>%
    mutate(tau = taui, islet = seq_along(phii))
}

# what fraction of contacts are on-island (vs. off?)
on_island_f = Vectorize(function(n, a) {
  d = floor(n / 2) - 1
  (exp(a) - 1) * exp(a * d) / (exp(a * d) + exp(a * (d + 1)) - 2)
})

# find the alpha matching some on-island fraction
f_to_alpha = Vectorize(function(oif, n) {
  uniroot(function(a) on_island_f(n, a) - oif, c(1e-12, 10))$root
})

n_ring = 50
ring_tau = (3.0 / 12) / floor(n_ring / 2) * ring_dist(1, 1:n_ring, n_ring)

rings = data_frame(on_island_f = c(0.99, 0.50, 0.25, 0.10, 0.05)) %>%
  mutate(alpha = f_to_alpha(on_island_f, n_ring)) %>%
  mutate(model = map(alpha, ~ ring_sim(ring_tau, alpha = .))) %>%
  unnest() %>%
  mutate(rho = R / (S + R + D))
```

```{r ring_plot}
# compare with within-host neutral cross-island model
# black lines
rings %>%
  ggplot(aes(tau, rho)) +
  geom_line(data = filter(davies_cross, tau <= 0.25)) +
  geom_line(data = filter(davies_small, big_tau == 0.125), linetype = 2) +
  geom_point(aes(color = factor(on_island_f))) +
  geom_line(aes(color = factor(on_island_f)))
```
The colored points and lines show the ring simulations. The solid black line shows the original cross-island simulation. The dotted black line shows a small-island simulation (i.e., fraction on island approximately zero) with a big-island $\tau$ equal to the mean of the ring-island $\tau$.

### A ring with randomly arranged use

What if islands had randomly mixed uses? Like, there's no spatial autocorrelation: each use is drawn from a random uniform distribution.

Now, a larger on-island proportion is required to see a result.

```{r random_ring}
ring_tau2 = runif(n_ring, 0, 3.0 / 12)

rings2 = data_frame(on_island_f = c(0.99, 0.50, 0.25, 0.10, 0.05)) %>%
  mutate(alpha = f_to_alpha(on_island_f, n_ring)) %>%
  mutate(model = map(alpha, ~ ring_sim(ring_tau2, alpha = .))) %>%
  unnest() %>%
  mutate(rho = R / (S + R + D))

rings2 %>%
  ggplot(aes(tau, rho)) +
  geom_line(data = davies_cross) +
  geom_line(data = filter(davies_small, big_tau == 0.125), linetype = 2) +
  geom_point(aes(color = factor(on_island_f))) +
  geom_smooth(aes(color = factor(on_island_f)), se = FALSE, size = 0.5)
```
