---
title: "Geography and associations between antibiotic use and resistance: simulations"
author: "Scott Olesen"
---

```{r global_options, echo = FALSE, message = FALSE}
library(knitr)
library(cowplot)
library(gridExtra)
library(tidyverse)

pdf.options(useDingbats = FALSE, useKerning = FALSE)

knitr::opts_chunk$set(
  fig.width = 7, fig.height = 4, fig.path = 'fig/',
  dev = c('png', 'pdf'),
  echo = FALSE, warning = FALSE, message = FALSE,
  cache = TRUE, autodep = TRUE)
```

```{r helpers}
annotate_text <- function(x, y, label, parse = TRUE) {
  annotate('text', x = x, y = y, label = label, parse = parse)
}
```

# Diagram

```{r two_pop_diagram}
dummy_data = tibble(
  x = c(-2, 2),
  y = c(-2, 2)
)

circle_data_f = function(cx, cy, r, n = 1e2) {
  tibble(
    theta = seq(0, 2 * pi, length.out = n),
    x = cx + r * cos(theta),
    y = cy + r * sin(theta)
  )
}

circle_data = bind_rows(
  circle_data_f(-1, 0, 0.5) %>% mutate(fill = 'white'),
  circle_data_f(1, 0, 0.5) %>% mutate(fill = 'black')
)

line_data = tibble(
  x = c(-1, -1, 1, 0),
  xend = c(1, -1, 1, 0),
  y = 1.0 * c(0, 1, 1, -1),
  yend = c(0, 0, 0, 0),
  linetype = c('a', rep('b', 3))
)

twopop_diagram <- ggplot(data = NULL, aes(x, y)) +
  geom_blank(data = dummy_data) +
  geom_segment(
    data = line_data,
    aes(xend = xend, yend = yend, linetype = linetype)
  ) +
  annotate_text(-1, 1.4, 'intervention\npopulation', parse = FALSE) +
  annotate_text(1, 1.4, 'control\npopulation', parse = FALSE) +
  annotate_text(
    0, -1.4,
    "paste('interaction strength (', epsilon, ')', )"
  ) +
  geom_polygon(data = circle_data, aes(fill = fill), color = 'black') +
  scale_fill_manual(values = c('white', 'black')) +
  guides(fill = 'none', linetype = 'none') +
  coord_fixed() +
  theme_void()

twopop_diagram
```

# Within-host neutral (WHN) model

```{r whn_parameters}
whn_base_parms <- list(
  beta = 4.0,
  u = 1.0,
  cost = 0.1683543
)

whn_base_parms %>%
  as_tibble() %>%
  kable(caption = 'WHN model parameters')
```

## WHN single-population model

```{r whn_sim}
# The ODE function only takes a vector, so we need to "unpack" the
# vector into a sensible data frame, compute the ODEs, then "pack"
# the data frame back into a vector.
whn_unpack <- function(x) {
  matrix(x, ncol = 4) %>%
    as_tibble() %>%
    setNames(c('X', 'S', 'R', 'D'))
}

# data frame to vector
whn_pack <- function(df) {
  stopifnot(names(df) == c('X', 'S', 'R', 'D'))
  unlist(df)
}

whn_ode_func <- function(t, state_vector, parms) {
  state <- whn_unpack(state_vector)

  with(c(state, parms), {
    n_pop <- nrow(state) # number of populations
    stopifnot(dim(betaij) == rep(n_pop, 2)) # betaij is defined in the next function

    N <- X + S + R + D # total number of individuals

    dS <- (betaij %*% ((S + D) / N)) * X - (u + taui) * S - (1 - cost) * (betaij %*% (R / N)) * S
    dR <- (1 - cost) * (betaij %*% (R / N)) * X - u * R + taui * D
    dD <- (1 - cost) * (betaij %*% (R / N)) * S - (u + taui) * D
    dX <- -(dS + dR + dD)

    list(c(dX, dS, dR, dD))
  })
}

# Check that:
# - All entries in the transmission matrix are nonnegative
# - All rows add up to the overall constant
# - All columns add up to the overall constant
check_beta <- function(mat, const) {
  stopifnot(all(mat >= 0))
  stopifnot(all(rowSums(mat) == const))
  stopifnot(all(colSums(mat) == const))
}

# Run the simulation
# taui: a vector of the antibiotic use rates in each population
whn_sim <- function(taui, epsilon) {
  n_pop <- length(taui)
  stopifnot(between(epsilon, 0, 1 - 1 / n_pop))

  parms <- with(whn_base_parms, {
    # Compare to the equations in the Supplement under "WHN model with 2 populations"
    betaij = matrix(beta * epsilon, ncol = n_pop, nrow = n_pop)
    diag(betaij) <- beta * (1 - epsilon)

    check_beta(betaij, beta)

    # Add transmission matrix beta_ij to the parameter list
    c(whn_base_parms, list(
      taui = taui, epsilon = epsilon,
      betaij = betaij
    ))
  })

  # Initial state of the simulation
  state <- tibble(
    X = rep(0.990 / n_pop, n_pop),
    S = rep(0.005 / n_pop, n_pop),
    R = S,
    D = 0
  )

  state_vector <- whn_pack(state)

  # Run the simulation
  result <- rootSolve::runsteady(state_vector, func = whn_ode_func, parms = parms)

  # Unpack the output state and clean it up
  result$y %>%
    whn_unpack() %>%
    mutate(
      tau = taui,
      pop = seq_along(taui),
      rho = R / (S + R + D)
    )
}
```

To make sure I got the equations right, I reproduce the $k = 1$ line in Davies *et al.* Figure 2g.

```{r davies_repro}
whn_repro <- whn_sim(taui = seq(0, 3 / 12, length.out = 16), epsilon = 0.0)

whn_repro %>%
  ggplot(aes(tau, rho)) +
  geom_point() +
  geom_line() +
  scale_x_continuous(
    name = 'antibiotic tx per month',
    sec.axis = sec_axis(~ . * 12)
  ) +
  ylab('proportion resistant') +
  theme_cowplot()
```

## WHN two-population model

```{r whn_double}
whn_epsilon_values <- c(0, 1e-4, 1e-3, 0.01, 0.0175, 0.025, 0.050, 0.075, 0.1, 0.2, 0.3, 0.4, 0.5)

whn2 <- tibble(
  delta_tau = round(seq(0.0, 0.15, length.out = 7), 3),
  taui = map(delta_tau, ~ 0.125 + c(-0.5, 0.5) * .)
) %>%
  crossing(epsilon = whn_epsilon_values) %>%
  mutate(
    results = map2(taui, epsilon, whn_sim),
    delta_rho = map_dbl(results, ~ max(.$rho) - min(.$rho))
  )
```

```{r whn_barchart_data}
whn2_plot_data <- whn2 %>%
  mutate(pop = map(1:n(), ~ c('intervention', 'control'))) %>%
  unnest() %>%
  filter(epsilon %in% c(0.0, 0.01, 0.1))
```

```{r whn_barcharts}
whn2_plot_f <- function(df) {
  initial_resistance <- df %>%
    filter(epsilon == 0, pop == 'control') %>%
    pull(rho)

  df %>%
    ggplot(aes(x = factor(epsilon), y = rho, fill = pop)) +
    geom_hline(yintercept = initial_resistance, linetype = 2) +
    geom_col(position = 'dodge', color = 'black') +
    scale_fill_manual(
      '',
      values = c('white', 'black')
    ) +
    scale_x_discrete(name = expression(epsilon)) +
    scale_y_continuous(
      name = 'resistance (%)',
      labels = scales::percent_format(accuracy = 1, suffix = ''),
      limits = c(0, 1.0),
      expand = c(0, 0)
    ) +
    theme_half_open(font_size = 12) +
    theme(
      legend.key.size = unit(3, 'mm'),
      legend.position = c(0.3, 1.03),
      axis.ticks.x = element_blank()
    )
}

whn2_plot1 <- whn2_plot_data %>%
  filter(delta_tau == 0.05) %>%
  whn2_plot_f()

whn2_plot2 <- whn2_plot_data %>%
  filter(delta_tau == 0.10) %>%
  whn2_plot_f()
```

```{r whn_dr_vs_dtau}
whn2_plot3 <- whn2 %>%
  filter(epsilon %in% c(0.0, 0.01, 0.1)) %>%
  ggplot(aes(delta_tau, delta_rho, group = factor(epsilon))) +
  geom_point() +
  geom_line() +
  scale_x_continuous(
    expression(Delta * tau),
    limits = c(0, 0.15),
    expand = c(0, 0.002)
  ) +
  scale_y_continuous(
    expression(Delta * rho),
    limits = c(0, 0.80),
    expand = c(0, 0.01)
  ) +
  annotate_text(0.07, 0.60, "epsilon == 0") +
  annotate_text(0.10, 0.35, "epsilon == 0.01") +
  annotate_text(0.13, 0.12, "epsilon == 0.1") +
  theme_cowplot(font_size = 12)

show(whn2_plot3)
```

```{r whn_drdtau_vs_epsilon}
whn2_plot4 <- whn2 %>%
  filter(
    delta_tau %in% c(0.05, 0.10),
    epsilon %in% c(0, 1e-4, 0.01, 0.025, 0.050, 0.075, 0.1, 0.2, 0.3, 0.4, 0.5)
  ) %>%
  mutate(dr_du = delta_rho / delta_tau) %>%
  ggplot(aes(epsilon, dr_du, group = factor(delta_tau))) +
  geom_point(aes(shape = factor(delta_tau))) +
  geom_line() +
  geom_blank(data = tibble(epsilon = 0, dr_du = 0, delta_tau = 0.01)) +
  scale_shape_manual(
    values = c(1, 16),
    labels = c(
      expression(Delta * tau == 0.05),
      expression(Delta * tau == 0.10)
    )
  ) +
  guides(shape = guide_legend(title = "", label.hjust = 0)) +
  scale_x_continuous(
    expression(epsilon),
    expand = c(0, 0.005)
  ) +
  scale_y_continuous(
    expression(Delta * rho / Delta * tau),
    limits = c(0, 7),
    expand = c(0, 0.1)
  ) +
  theme_cowplot(font_size = 12) +
  theme(legend.position = c(0.5, 0.75))

whn2_plot4
```

```{r whn_double_results}
whn2_plot_nolabel <- gridExtra::grid.arrange(
  grobs = list(
    twopop_diagram, whn2_plot1, whn2_plot2,
    whn2_plot3, whn2_plot4
  ),
  layout_matrix = rbind(rep(1:3, each = 2), rep(4:5, each = 3))
)

whn2_plot <- ggdraw(whn2_plot_nolabel) +
  annotate_text(0.015, 0.98, 'bold("a)")') +
  annotate_text(0.350, 0.98, 'bold("b)")') +
  annotate_text(0.680, 0.98, 'bold("c)")') +
  annotate_text(0.015, 0.48, 'bold("d)")') +
  annotate_text(0.550, 0.48, 'bold("e)")')

ggsave(
  'fig/whn2_plot.pdf', whn2_plot,
  width = 190, height = 100, unit = 'mm'
)

ggsave(
  'fig/whn2_plot.png', whn2_plot,
  width = 190, height = 100, unit = 'mm'
)
```

# D-types model

```{r dtypes_parameters}
dtypes_base_parms <- list(
  n_d = 16,
  min_mu = 0.5,
  max_mu = 2.0,
  beta = 2.0,
  c_beta = 1.0,
  c_mu = 1.1,
  k = 15.0
)

dtypes_base_parms %>%
  as_tibble() %>%
  kable(caption = 'D-types base parameters')

dtypes_base_parms$mu_d <- with(
  dtypes_base_parms,
  seq(min_mu, max_mu, length = n_d)
)

with(
  dtypes_base_parms,
  tibble(
    d = 1:n_d,
    mu_d = mu_d
  )
) %>%
  kable(caption = 'D-types mu_d values')
```

## Single-population model

```{r dtype_helpers}
# pack the list into a vector
dtypes_pack <- function(lst) {
  stopifnot(setequal(names(lst), c('X_i', 'S_id', 'R_id')))
  stopifnot(all(dim(lst$S_id) == dim(lst$R_id)))
  stopifnot(dim(lst$S_id)[1] == length(lst$X_i))
  c(lst$X_i, as.vector(lst$S_id), as.vector(lst$R_id))
}

# unpack the vector into a list
dtypes_unpack <- function(x, n_pop, n_d) {
  stopifnot(length(x) == n_pop + 2 * (n_d * n_pop))

  x_end <- n_pop
  s_start <- x_end + 1
  s_end <- x_end + n_d * n_pop
  r_start <- s_end + 1
  r_end <- s_end + n_d * n_pop
  stopifnot(r_end == length(x))

  list(
    X_i = x[1:x_end],
    S_id = matrix(x[s_start:s_end], nrow = n_pop, ncol = n_d),
    R_id = matrix(x[r_start:r_end], nrow = n_pop, ncol = n_d)
  )
}

# for initially packing the data frame into the list
dtypes_df_to_list <- function(df) {
  # check we have the right column names
  stopifnot(setequal(names(df), c('population', 'phenotype', 'dtype', 'value')))
  # check we have the right phenotypes
  stopifnot(setequal(df$phenotype, c('S', 'R', 'X')))

  # extract the no. populations and D-types
  n_pop <- length(unique(df$population))
  n_d <- length(unique(df$dtype))

  # there should be S and R for each pop/D-type combo
  stopifnot(nrow(filter(df, phenotype == 'S')) == n_pop * n_d)
  stopifnot(nrow(filter(df, phenotype == 'R')) == n_pop * n_d)
  # but X for only each pop
  stopifnot(nrow(filter(df, phenotype == 'X')) == n_pop)

  list(
    X_i = df %>% filter(phenotype == 'X') %>% pull(value),
    S_id = df %>% filter(phenotype == 'S') %>% pull(value) %>% matrix(nrow = n_pop, ncol = n_d),
    R_id = df %>% filter(phenotype == 'R') %>% pull(value) %>% matrix(nrow = n_pop, ncol = n_d)
  )
}

# for finally unpacking the list into a data frame
dtypes_list_to_df <- function(lst) {
  # check for names
  stopifnot(setequal(names(lst), c('X_i', 'S_id', 'R_id')))

  # get dimensions
  n_pop <- length(lst$X_i)
  n_d <- dim(lst$S_id)[2]

  # check shapes
  stopifnot(all(dim(lst$S_id) == c(n_pop, n_d)))
  stopifnot(all(dim(lst$R_id) == c(n_pop, n_d)))

  X_rows <- tibble(pop = 1:n_pop, phenotype = 'X', dtype = NA, value = lst$X_i)
  S_rows <- crossing(dtype = 1:n_d, pop = 1:n_pop) %>%
    mutate(phenotype = 'S', value = as.vector(lst$S_id))
  R_rows <- crossing(dtype = 1:n_d, pop = 1:n_pop) %>%
    mutate(phenotype = 'R', value = as.vector(lst$R_id))

  bind_rows(X_rows, S_rows, R_rows) %>%
    arrange(pop, phenotype, dtype)
}
```

```{r dtypes_sim}
dtypes_ode_func <- function(t, state_vector, parms) {
  state <- dtypes_unpack(state_vector, n_pop = parms$n_pop, n_d = parms$n_d)

  with(c(state, parms), {
    stopifnot(length(X_i) == n_pop)
    stopifnot(dim(S_id) == c(n_pop, n_d))
    stopifnot(dim(R_id) == c(n_pop, n_d))

    # rowSums are over D-types (second index)
    v_id <- (1.0 - ((S_id + R_id) / rowSums(S_id + R_id) - 1.0 / n_d)) ** k
    stopifnot(dim(v_id) == c(n_pop, n_d))

    N_i <- X_i + rowSums(S_id + R_id)
    stopifnot(length(N_i) == n_pop)

    # "tau_i * S_id" does sum_i { tau_i S_id }, which is length P vector
    # to multiply by rows, need to do some fancy footwork: "mat %*% diag(row)"
    dS_id <- v_id * (beta_ij %*% (S_id / N_i)) * X_i - tau_i * S_id - S_id %*% diag(mu_d)
    dR_id <- v_id * (beta_ij %*% (R_id / N_i)) * X_i - c_mu * R_id %*% diag(mu_d)
    dX_i <- -rowSums(dS_id + dR_id)

    list(dtypes_pack(list(X_i = dX_i, S_id = dS_id, R_id = dR_id)))
  })
}

dtypes_sim <- function(tau_i, epsilon = 0) {
  n_pop <- length(tau_i)
  stopifnot(between(epsilon, 0, 1 - 1 / n_pop))

  parms <- with(dtypes_base_parms, {
    beta_ij = matrix(beta * epsilon, ncol = n_pop, nrow = n_pop)
    diag(beta_ij) <- beta * (1 - epsilon)

    check_beta(beta_ij, beta)

    c(dtypes_base_parms, list(
      n_pop = n_pop, tau_i = tau_i, epsilon = epsilon, beta_ij = beta_ij
    ))
  })

  n_d <- parms$n_d

  state <- list(
    X_i = rep(0.9 / n_pop, n_pop),
    S_id = matrix(0.05 / (n_pop * n_d), nrow = n_pop, ncol = n_d),
    R_id = matrix(0.05 / (n_pop * n_d), nrow = n_pop, ncol = n_d)
  )

  state_vector <- dtypes_pack(state)
  stopifnot(all.equal(sum(state_vector), 1))

  result <- rootSolve::runsteady(
    state_vector,
    func = dtypes_ode_func,
    parms = parms,
    stol = 1e-8 / n_pop,
    rtol = 1e-6 / n_pop,
    atol = 1e-6 / n_pop
  )

  result$y %>%
    dtypes_unpack(n_pop = n_pop, n_d = n_d) %>%
    dtypes_list_to_df() %>%
    left_join(tibble(pop = 1:n_pop, tau = tau_i), by = 'pop')
}
```

To ensure I got the equations right, I reproduce Figure 3 with $\tau = 0.075$.

```{r dtypes_repro_bars}
dtype_repro_bars <- dtypes_sim(0.075, 0)

dtype_repro_bars %>%
  mutate(mu_d = dtypes_base_parms$mu_d[dtype]) %>%
  ggplot(aes(mu_d, value, fill = phenotype)) +
  geom_col() +
  xlab('Clearance rate') +
  ylab('Strain prevalence') +
  theme_cowplot()
```

Next, I show the jagged line in Figure 4. It may not look exactly right because I ran a limited number of simulations:

```{r dtypes_repro_line}
dtype_repro_line <- dtypes_sim(seq(0.04, 0.20, length.out = 50), 0)

dtype_repro_line %>%
  group_by(tau) %>%
  summarize(
    f_res = sum(value[phenotype == 'R']),
    f_all = sum(value)
  ) %>%
  mutate(rho = f_res / f_all) %>%
  ggplot(aes(tau, rho)) +
  geom_point() +
  geom_line() +
  theme_cowplot()
```

## D-types two-population

```{r dtypes_double}

dtypes_epsilon_values <- c(0, 1e-4, 0.001, 0.005, 0.01, 0.0175, 0.025, 0.05, 0.075, 0.1)

dtypes_simplify_results <- function(df) {
  df %>%
    group_by(pop, phenotype) %>%
    summarize(
      value = sum(value),
      tau = unique(tau)
    ) %>%
    ungroup() %>%
    spread(phenotype, value) %>%
    mutate(rho = R / (R + S))
}

dtypes2 <- tibble(
  delta_tau = c(0.05, 0.10),
  taui = map(delta_tau, ~ 0.125 + c(-0.5, 0.5) * .)
) %>%
  crossing(epsilon = dtypes_epsilon_values) %>%
  mutate(
    raw_results = map2(taui, epsilon, dtypes_sim),
    results = map(raw_results, dtypes_simplify_results),
    delta_rho = map_dbl(results, ~ max(.$rho) - min(.$rho)),
    dr_du = delta_rho / delta_tau
  )
```

```{r dtypes_plot}
dtypes2_plot <- dtypes2 %>%
  ggplot(aes(epsilon, dr_du, group = factor(delta_tau))) +
  geom_point(aes(shape = factor(delta_tau))) +
  geom_line() +
  geom_blank(data = tibble(epsilon = 0, dr_du = 0, delta_tau = 0.01)) +
  scale_shape_manual(
    values = c(1, 16),
    labels = c(
      expression(Delta * tau == 0.05),
      expression(Delta * tau == 0.10)
    )
  ) +
  guides(shape = guide_legend(title = "", label.hjust = 0)) +
  scale_x_continuous(
    expression(epsilon),
    limits = c(0, 0.1),
    expand = c(0.02, 0, 0.05, 0)
  ) +
  scale_y_continuous(
    expression(Delta * rho / Delta * tau),
    limits = c(0, 6.5),
    expand = c(0, 0.1)
  ) +
  theme_cowplot(font_size = 12) +
  theme(legend.position = c(0.5, 0.75))

show(dtypes2_plot)

ggsave(
  'fig/dtypes2_plot.pdf', dtypes2_plot,
  width = 88, height = 80, unit = 'mm'
)

ggsave(
  'fig/dtypes2_plot.png', dtypes2_plot,
  width = 88, height = 80, unit = 'mm'
)
```

# Reduction in $\Delta \rho / \Delta \tau$ against $\epsilon$

```{r reduction}
bind_rows(
  'WHN' = whn2,
  'Dtypes' = dtypes2,
  .id = 'model'
) %>%
  mutate(
    dr_du = delta_rho / delta_tau,
    key = str_c(model, ' Δτ=', delta_tau)
  ) %>%
  filter(
    delta_tau %in% c(0.05, 0.10),
    epsilon %in% c(0, 1e-4, 1e-3, 1e-2, 0.0175, 0.025)
  ) %>%
  group_by(model, delta_tau) %>%
  mutate(pct_red = scales::percent(1 - dr_du / dr_du[epsilon == 0])) %>%
  ungroup() %>%
  select(epsilon, key, pct_red) %>%
  spread(key, pct_red) %>%
  kable(caption = 'Percent reduction, compared to ε=0')
```
